### 2018-09-06 paper.Rmd
updates

---
title: "Football Championship Ticket Options"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Since its adoption in the 2014-2015 season, the College Football Playoff (CFP) has given college football fans even more reason to be excited about rankings. No longer did a team need to be in the top 2 of the Bowl Championship Series (BCS) to be a championship contender -- instead, the top 4 teams entered into a playoff. In capitalizing on the intersection between excitement and fandom, the CFP teamed with OptionIt to introduce an RSVP system for championship game tickets, starting in 2018. A fan may by an RSVP (an option) for a team. If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value. In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market.

A source of inquiry is what drives the price of these RSVPs. Before the season begins, would knowing a teams performance during the previous year drive prices? Would preseason ranking have any influence on prices? After the start of the season, would rankings, performance, injuries, or other variables affect prices? These questions will be explored throughout our paper.

# Hypotheses

# Methods

## Data

```{r}
library(dplyr)

optionsDatNew = readr::read_csv("optionsDatNew.csv")

test = optionsDatNew %>% 
  select(name, dateTime, contains("price"))

library(ggplot2)

alabamaTS = test %>% 
  filter(name == "Alabama") %>% 
  select(dateTime, LL_price) %>% 
  mutate(dateTime = lubridate::ymd_hms(dateTime), 
         dateTime = lubridate::day(dateTime)) %>% 
  ts(.)

library(forecast)

alabamaForecast = HoltWinters(alabamaTS, beta = FALSE, gamma = FALSE)

```

### Sources

Given the research questions.

### Issues and Limitations

# Results

### Previous Year-end Rank and Initial Prices

```{r}
initialData = readr::read_csv("optionsDat.csv")

previousYearRank = readr::read_csv("massey2017YearEnd.csv")

initialData = initialData %>% 
  left_join(., previousYearRank, by = c("name" = "Team")) %>% 
  group_by(name)

initialData %>% 
  slice(., 1L) %>% 
  filter(masseyYearEnd < 50) %>% 
  ggplot(., aes(masseyYearEnd, as.numeric(LL_price))) +
  geom_point()

```

### 2018-09-06 masseyPreviousYear.R

@@ -1,0 +1,22 @@
+############################
+### Massey Previous Year ###
+############################
+
+library(jsonlite)
+
+masseyPage = fromJSON("https://www.masseyratings.com/teamjson.php?t=11604&s=295489", 
+                      flatten = FALSE)
+
+columnNames = masseyPage$CI$title
+
+ratingsTable = matrix(unlist(masseyPage$DI), nrow = length(masseyPage$DI), 
+                      ncol = length(unlist(masseyPage$DI[[1]])), 
+                      byrow = TRUE) %>% 
+  data.frame()
+
+ratingsTable = select(ratingsTable, 
+                      X1, X2)
+
+names(ratingsTable) = c("masseyYearEnd", "Team")
+
+write.csv(ratingsTable, "massey2017YearEnd.csv", row.names = FALSE)

### 2018-09-06 masseyRatings.R

@@ -35,5 +35,3 @@ masseyRatings = function() {
   return(ratingsTable)
   
 }
-
-masseyRatings()

### 2018-09-06 teamOptionsPullPrep.R

@@ -1,8 +1,28 @@
 ##################################
-### Team Options Pull and Prep ###
+###       Team Options         ### 
+###     Pull, Prep, Write      ###
 ##################################
 
+# The following packages will be needed for everything within this file
+# and the sourced functions.
+
+if (!require("pacman")) install.packages("pacman")
+
+pacman::p_load(dplyr, jsonlite, tidyr, readr, rvest, taskscheduleR)
+
 library(dplyr); library(jsonlite); library(tidyr)
+
+# Sourcing the following files will run the appropriate functions.
+# Before sourcing, however, you will need to specify the appropriate locations
+# for your folder structure. Uncomment the code on lines 20 through 24 to run
+# everything on your own machine.
+
+# pathToMyFolder = "path/to/files/"
+# source(paste(pathToMyFolder, "injuryScrape.R", sep = "))
+# source(paste(pathToMyFolder, "masseyComps.R", sep  = "))
+# source(paste(pathToMyFolder, "masseyRatings.R", sep  = "))
+# source(paste(pathToMyFolder, "predictedWinners.R", sep = "))
+
 
 source("C:/Users/sberry5/Documents/research/ticketOptions/injuryScrape.R")
 
@@ -12,32 +32,19 @@ source("C:/Users/sberry5/Documents/research/ticketOptions/masseyRatings.R")
 
 source("C:/Users/sberry5/Documents/research/ticketOptions/predictedWinners.R")
 
-# Thankfully, we have this really nice json file!
 
-# optionsDat = jsonlite::read_json("https://api.cfp-rsvp.com/api/teams/getTeams?eventId=2", 
-#                            simplifyVector = TRUE)
+# First, we will get the options data from CFP-RSVP. The information that we want
+# is sitting a json file.
 
 optionsDat = jsonlite::read_json("https://api.dibitnow.com/api/v1/eventPerformers/getEventPerformers?eventId=1", 
                                  simplifyVector = TRUE)
 
-# To keep track of date and time, we will add Sys.time. 
+# To keep track of date and time for each pull, we will add Sys.time. 
 
 optionsDat$dateTime = Sys.time()
 
 # One of the columns, teamTiers, has a nested list structure. We need to spread
 # it out, collapse it down, and then bind the individual data frames.
-
-# spreadData = lapply(1:nrow(optionsDat), function(x) {
-#   
[…]

### 2018-09-07 paper.Rmd
Data reconcile and updates

@@
 
-```{r}
-library(dplyr)
-
-optionsDatNew = readr::read_csv("optionsDatNew.csv")
-
-test = optionsDatNew %>% 
-  select(name, dateTime, contains("price"))
-
-library(ggplot2)
-
-alabamaTS = test %>% 
-  filter(name == "Alabama") %>% 
-  select(dateTime, LL_price) %>% 
-  mutate(dateTime = lubridate::ymd_hms(dateTime), 
-         dateTime = lubridate::day(dateTime)) %>% 
-  ts(.)
-
-library(forecast)
-
-alabamaForecast = HoltWinters(alabamaTS, beta = FALSE, gamma = FALSE)
-
-```
-
 ### Sources
@@
 
+As a starting point, we began by looking at Massey's year-end rankings and early option prices (February 2018).
+
+In an effort to reduce visual clutter and focus on potentially high-value teams, we can visually inspect the top 25 teams at the end of the 2017-2018 season.
+
 ```{r}
@@
 
+previousYearRank = previousYearRank %>% 
+  mutate(Team = as.character(Team), 
+         Team = gsub("St$", "State", Team))
+
 initialData = initialData %>% 
+  mutate(name = gsub("\\.|,", "", name)) %>% 
   left_join(., previousYearRank, by = c("name" = "Team")) %>% 
@@
   slice(., 1L) %>% 
-  filter(masseyYearEnd < 50) %>% 
-  ggplot(., aes(masseyYearEnd, as.numeric(LL_price))) +
-  geom_point()
+  filter(masseyYearEnd < 25) %>% 
+  ggplot(., aes(masseyYearEnd, as.numeric(LL_price), color = as.factor(masseyYearEnd))) +
+  geom_point() +
+  theme_minimal()
+```
+
+Visually, there is cer
+
+### Prices By Conference
+
+```{r}
 

### 2018-09-07 dataReconcile.R

@@ -1,0 +1,29 @@
+########################################
+### Data Version and Reconcilliation ###
+########################################
+
+library(dplyr)
+
+# Given the various computer and data source changes, we have a few data files
+# that need checked and bound together.
+
+originalData = readr::read_csv("optionsDat.csv")
+
+# Since the api changed, we need to select key variables, 
+# so that we can get them bound to subsequent files.
+
+originalData = originalData %>% 
+  select(name, dateTime, initialPrice, price, LL_maxPurchaseQuantity, LL_price, 
+         UL_maxPurchaseQuantity, UL_price)
+
+
+updatedData = readr::read_csv("optionsDatNew.csv")
+
+updateV2Data = readr::read_csv("optionsDatV2.csv")
+
+finalData = bind_rows(updateV2Data, updatedData, originalData)
+
+finalData = finalData %>% 
+  mutate(name = gsub("\\.|,", "", name))
+
+write.csv(finalData, "completeData.csv", row.names = FALSE)

### 2018-09-07 teamOptionsPullPrep.R

@@ -75,13 +75,24 @@ optionsDat = cbind(optionsDat, spreadData) %>%
   select(-eventPerformerCategory, -eventPerformerExperiences, 
          -priceTrend)
 
+optionsDat = optionsDat %>% 
+  mutate(name = gsub("\\.|,", "", name))
+
 # Now, we can use our sourced functions and finally join everything together.
 
 injury = injuryScrape()
 
 masseyCompsDat = masseyComps()
 
+masseyCompsDat = masseyCompsDat %>% 
+  mutate(Team = as.character(Team), 
+         Team = gsub("St$", "State", Team))
+
 masseyRatingsDat = masseyRatings()
+
+masseyRatingsDat = masseyRatingsDat %>% 
+  mutate(Team = as.character(Team), 
+         Team = gsub("St$", "State", Team))
 
 winners = predictedWinners()
 
@@ -96,5 +107,5 @@ optionsDat = optionsDat %>%
 # the same file.
 
 write.table(x = optionsDat, 
-            file = "C:/Users/sberry5/Documents/research/ticketOptions/optionsDatV2.csv", append = TRUE, 
+            file = "C:/Users/sberry5/Documents/research/ticketOptions/completeData.csv", append = TRUE, 
             na = "", sep = ",", row.names = FALSE)

### 2018-09-13 paper.Rmd
updates

@@
 In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market.
+In financial modeling terms, this system could be considered a call option for speculation.
+When buying an option (the RSVP), the buyer has the right to buy a stock (a ticket) in the future at the strike price (the face value of the ticket).
 
@@
 
+The RSVP data presented a few challenges along the way -- mainly that the data source changed at some point during the course of data collection.
+This data source change resulted in a considerable span of data loss.
+Further issues arised on the automatic nature of data collection.
+Windows Task Scheduler was used to automatically run our data collection scripts daily -- unfortunately, this will not run if a computer is shut down.
+
 # Results
 
-### Previous Year-end Rank and Initial Prices
+### Previous Year-end Rank and Early Prices
 
@@
 ```{r}
+
+library(dplyr)
+
+library(ggplot2)
+
 initialData = readr::read_csv("optionsDat.csv")
@@
   slice(., 1L) %>% 
-  filter(masseyYearEnd < 25) %>% 
-  ggplot(., aes(masseyYearEnd, as.numeric(LL_price), color = as.factor(masseyYearEnd))) +
-  geom_point() +
+  dplyr::filter(masseyYearEnd < 26) %>% 
+  select(masseyYearEnd, UL_price, LL_price, name) %>% 
+  tidyr::gather(key, value, -name, -masseyYearEnd) %>% 
+  arrange(name) %>% 
+  ggplot(., aes(name, as.numeric(value), color = key)) +
+  geom_point(size = 3) +
+  scale_x_discrete(limits = top25PrevNames$Team) +
+  theme_minimal()
+
+initialData %>% 
+  slice(., 1L) %>% 
+  select(name, masseyYearEnd, UL_price, LL_price) %>% 
+  arrange(masseyYearEnd, UL_price)
+```
+
+While North Dakota State finished last year at 23 according to Massey, they are not part of the RSVP data.
+
+## Replicating Rankings Through Option Prices
+
+### Top 10 Trends
+
+The following figure shows the price trend for every team that has been in the AP top 10 since the beginning of the season.
+
+```{r}
+completeData = read_csv("completeData.csv")
+
+top10Ever = unique(completeData$name[as.numeric(completeData$AP) < 11])
+
+endOfDay = completeData %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
+  group_by(name, day) %>% 
+  slice(n())
+
+completeData %>% 
+  mutate(Rank = as.numeric(Rank), 
+         dateTime = lubridate::ymd_hms(dateTime)) %>% 
+  dplyr::filter(name %in% top10Ever) %>% 
+  select(dateTime, UL_price, LL_price, name) %>% 
+  tidyr::gather(key, value, -name, -dateTime) %>% 
+  arrange(name, dateTime) %>% 
+  ggplot(., aes(dateTime, as.numeric(value), color = key, group = key)) + 
+  geom_path(na.rm = TRUE, linejoin = "mitre") + 
+  facet_wrap(~ name) +
   theme_minimal()
@@
 
-Visually, there is cer
-
-### Prices By Conference
+###
 
 ```{r}
+library(lme4)
 
+top10 = completeData %>% 
+  dplyr::filter(name %in% top10Ever) %>% 
+  mutate(Rank = as.numeric(Rank), 
+         dateTime = lubridate::ymd_hms(dateTime), 
+         averagePrice = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
+  select(name, dateTime, averagePrice, Rank, AP, SAG, MAS, USA, 
+         Mean, n, computerPick, publicPick) %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
+  group_by(name, day) %>% 
+  slice(n())
+
+mixedMod = lmer(averagePrice ~ as.numeric(Rank) + (1|name), data = top10)
 ```
+
+###
+
+```{r}
+library(RQuantLib)
+
+AmericanOption(type = "call", 
+               underlying = , 
+               strike = 475, 
+               )
+```

### 2018-09-13 injuryScrape.R

@@ -30,7 +30,7 @@ injuryScrape = function() {
   test$team[which(test$team == "" & test$Date != "")] = NA
   
   test = tidyr::fill(test, team, .direction = "down") %>% 
-    filter(Date != "Date", Date != "", 
+    dplyr::filter(Date != "Date", Date != "", 
            grepl("^[A-Z]", .$Date) != TRUE) %>% 
     group_by(team) %>% 
     summarize(n = n())

### 2018-09-13 shuffle.R

@@ -1,0 +1,20 @@
+completeData = completeData %>% 
+  mutate(marker = ifelse(eventId == "eventId", seq(1:nrow(.)), NA)) %>% 
+  tidyr::fill(., marker, .direction = "down")
+
+
+markerNames = sort(unique(completeData$marker))
+
+refresh = lapply(markerNames, function(x) {
+  newDat = completeData %>% 
+    dplyr::filter(marker == x)
+  
+  names(newDat) = newDat[1, ]
+  
+  newDat = newDat[-1, ]
+  
+  return(newDat)
+  
+})
+
+refresh = data.table::rbindlist(refresh, fill = TRUE)

### 2018-09-13 teamOptionsPullPrep.R

@@ -97,11 +97,12 @@ masseyRatingsDat = masseyRatingsDat %>%
 winners = predictedWinners()
 
 optionsDat = optionsDat %>% 
-  left_join(., masseyCompsDat, by = c("name" = "Team")) %>% 
-  left_join(., masseyRatingsDat, by = c("name" = "Team")) %>% 
   left_join(., injury, by = c("name" = "team")) %>% 
   mutate(computerPick = ifelse(.$name %in% winners$computerPicks, 1, 0), 
-         publicPick = ifelse(.$name %in% winners$publicPicks, 1, 0))
+         publicPick = ifelse(.$name %in% winners$publicPicks, 1, 0)) %>% 
+  left_join(., masseyRatingsDat, by = c("name" = "Team")) %>% 
+  left_join(., masseyCompsDat, by = c("name" = "Team"))
+  
 
 # Since we are doing this continually, we will append new data into
 # the same file.
@@ -109,3 +110,8 @@ optionsDat = optionsDat %>%
 write.table(x = optionsDat, 
             file = "C:/Users/sberry5/Documents/research/ticketOptions/completeData.csv", append = TRUE, 
             na = "", sep = ",", row.names = FALSE)
+
+
+# completeData2 %>% 
+#   select(eventId:UL_price, Conf:X59, masseyConference:masseyEL, n, computerPick, publicPick) %>% 
+#   write.csv(., "tofix.csv", row.names = FALSE)

### 2018-09-14 paper.Rmd
updates

@@
 If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value.
-In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market.
+In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of XXXX for the 2018 Championship game.
 In financial modeling terms, this system could be considered a call option for speculation.
@@
 
+The internet is full of interesting data and finding it is never too hard.
+While finding interesting data does not present a challenge, finding data that is free to use can be tricky.
+In the context of data, free to use does not necessarily mean free in a monetary sense.
+Instead of a paywall, data is often locked behind terms of service (ToS).
+While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through Javascrip) or ToS.
+With the appropriate accumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS.
+In scraping our data, we not only read the ToS, but we also examined each site's robots.txt.
+Since we wanted to conduct our data collection in a programmatic fashion, we wanted to ensure that we were not going to be running afoul of restrictions placed within the robots.txt file.
+To that end, all the data that we collected is free to use -- in other words, no restrictions apply.
+
 ### Sources
 
-Given the research questions.
+Given the research questions, we sought to collect data from several different sources. Given the excellent compilation of team rankings, Massey's website was used to amass rankings.
+
+Injury counts were pulled from Don Best's website and projected winners were pulled from XXX.
+
+Finally, the RSVP data was pulled from the json object on the CFP-RSVP site.
+While the main page displays the current option price for an upper-level ticket (UL_price), the json object also contains prices for lower-level tickets (LL_price).
+There are also different hotel packages within the data, but our primary focus will be on the two ticket prices.
 
@@
 This data source change resulted in a considerable span of data loss.
-Further issues arised on the automatic nature of data collection.
+Further, issues arised on the automatic nature of data collection.
 Windows Task Scheduler was used to automatically run our data collection scripts daily -- unfortunately, this will not run if a computer is shut down.
+Finally, new ratings were added to the Massey Ratings page.
+Given the position of the Massey ratings in our data join, many columns became displaced from their original position.
+While this was fixed, some values were not recovered.
 
@@
 
+top25PrevNames = previousYearRank %>% 
+  arrange(as.numeric(masseyYearEnd)) %>% 
+  head(25) %>% 
+  select(Team)
+
 initialData %>% 
@@
   select(masseyYearEnd, UL_price, LL_price, name) %>% 
+  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
   tidyr::gather(key, value, -name, -masseyYearEnd) %>% 
@@
   theme_minimal()
-
-initialData %>% 
-  slice(., 1L) %>% 
-  select(name, masseyYearEnd, UL_price, LL_price) %>% 
-  arrange(masseyYearEnd, UL_price)
 ```
 
-While North Dakota State finished last year at 23 according to Massey, they are not part of the RSVP data.
+The preceding visualization provides a clear picture
+
+## Ranking Correlations
+
+For the sake of exploration and to limit the chances of future multicollinearity, we wanted to explore how some of the more well-known ranking were correlated:
+
+```{r}
+completeData %>% 
+  select(AP, SAG, MAS, USA, FPI) %>% 
+  na.omit() %>% 
+  cor() %>% 
+  corrplot::corrplot.mixed()
+```
 
@@
 ```{r}
-completeData = read_csv("completeData.csv")
+completeData = readr::read_csv("refreshed.csv") %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime))
 
@@
 
+A quick visual inspection of every team that has been in the top 10 this year might suggest persistent high confidence for Alabama, Clemson, and Georgia (but to a slightly less degree).
+Not surprisingly, Ohio State and Oklahoma are starting to trend upwards; meanwhile, Notre Dame has consistently high prices (relatively speaking), but on a potential downward trend.
+
 ###
@@
 top10 = completeData %>% 
-  dplyr::filter(name %in% top10Ever) %>% 
+  # dplyr::filter(name %in% top10Ever) %>% 
   mutate(Rank = as.numeric(Rank), 
@@
 
-mixedMod = lmer(averagePrice ~ as.numeric(Rank) + (1|name), data = top10)
+mixedMod = lmer(averagePrice ~ as.numeric(USA) + as.numeric(MAS) + 
+                  (1|name), data = top10)
+```
+
+### Volatility
+
+```{r}
+volatile = completeData %>% 
+  group_by(name, day) %>%
+  slice(n()) %>%
+  group_by(name) %>% 
+  mutate(lagLL = dplyr::lag(LL_price), 
+         lagUL = dplyr::lag(UL_price)) %>% 
+  mutate(volatilityLL = ((LL_price / lagLL) - 1), 
+         volatilityUL = ((UL_price / lagUL) - 1)) %>% 
+  summarize(sdLL = sd(volatilityLL, na.rm = TRUE), 
+            sdUL = sd(volatilityUL, na.rm = TRUE), 
+            volLL = sdLL * 15.937, 
+            volUL = sdUL * 15.937, 
+            meanLL = mean(LL_price, na.rm = TRUE), 
+            meanUL = mean(UL_price, na.rm = TRUE)) %>% 
+  arrange(desc(volLL), desc(volUL))
 ```

### 2018-09-14 shuffle.R

@@ -1,13 +1,38 @@
+######################
+### Data Wrangling ###
+######################
+
+# The following operations were conducted on the complete data
+# so that columns were re-aligned. Given the position of 
+# the Massey variables within the original data, new
+# Massey variables caused disruptions in alignment of 
+# existing columns. To that end, each pull needed to be
+# broken apart, renamed where appropriate, and bound
+# back together.
+
+library(dplyr)
+
+completeData = readr::read_csv("completeData_Deprecated.csv")
+
 completeData = completeData %>% 
   mutate(marker = ifelse(eventId == "eventId", seq(1:nrow(.)), NA)) %>% 
   tidyr::fill(., marker, .direction = "down")
 
-
 markerNames = sort(unique(completeData$marker))
 
 refresh = lapply(markerNames, function(x) {
+  
+  # browser()
   newDat = completeData %>% 
-    dplyr::filter(marker == x)
+    dplyr::filter(marker == x) %>% 
+    select(-marker)
+  
+  for(i in 1:length(newDat)) {
+    if(is.na(newDat[1, i])) {
+     newDat[1, i] = names(newDat)[i] 
+    }
+  }
+  
   
   names(newDat) = newDat[1, ]
   
@@ -17,4 +42,6 @@ refresh = lapply(markerNames, function(x) {
   
 })
 
-refresh = data.table::rbindlist(refresh, fill = TRUE)
+refresh = bind_rows(refresh)
+
+write.csv(refresh, "refreshed.csv", row.names = FALSE)

### 2018-09-14 code/functions/injuryScrape.R
structure updates with some codes updates

@@ -1,0 +1,40 @@
+#####################
+### Injury Scrape ###
+#####################
+
+injuryScrape = function() {
+  
+  library(rvest)
+  
+  injuryPage = read_html("http://www.donbest.com/ncaaf/injuries/")
+  
+  teamNames = html_nodes(injuryPage, "td.statistics_table_header") %>% 
+    html_text()
+  
+  columnNames = html_nodes(injuryPage, "th.statistics_cellrightborder") %>% 
+    html_text() %>% 
+    unique()
+  
+  test = html_table(injuryPage, fill = TRUE) %>% 
+    `[[`(2)
+  
+  test = test[-1, ] %>% 
+    select(X1:X5)
+  
+  names(test) = columnNames
+  
+  test$team = test$Status
+  
+  test$team[!(test$Date %in% teamNames)] = ""
+  
+  test$team[which(test$team == "" & test$Date != "")] = NA
+  
+  test = tidyr::fill(test, team, .direction = "down") %>% 
+    dplyr::filter(Date != "Date", Date != "", 
+           grepl("^[A-Z]", .$Date) != TRUE) %>% 
+    group_by(team) %>% 
+    summarize(n = n())
+  
+  return(test)
+  
+}

### 2018-09-14 code/functions/masseyComps.R

@@ -1,0 +1,23 @@
+##########################
+### Massey Comparisons ###
+##########################
+
+masseyComps = function() {
+  
+  skipLines = readLines("https://www.masseyratings.com/cf/compare.csv")
+  
+  allComps = readr::read_csv("https://www.masseyratings.com/cf/compare.csv", 
+                             skip = (which(grepl("^Team", skipLines)) - 1))
+  
+  return(allComps)
+  
+}
+
+# library(irr)
+# 
+# iccRatings = allComps %>% 
+#   select(ARG:SOR)
+# 
+# icc(iccRatings, model = "oneway")
+# 
+# kappam.fleiss(iccRatings)

### 2018-09-14 code/functions/masseyRatings.R

@@ -1,0 +1,37 @@
+######################
+### Massey Ratings ###
+######################
+
+masseyRatings = function() {
+  
+  library(jsonlite)
+  
+  masseyPage = fromJSON("https://www.masseyratings.com/ratejson.php?s=300937&sub=11604", 
+                        flatten = FALSE)
+  
+  columnNames = masseyPage$CI$title
+  
+  ratingsTable = matrix(unlist(masseyPage$DI), nrow = length(masseyPage$DI), 
+                        ncol = length(unlist(masseyPage$DI[[1]])), 
+                        byrow = TRUE) %>% 
+    data.frame()
+  
+  ratingsTable = select(ratingsTable, 
+                        -X2, -X3, 
+                        -X5, -X6)
+  
+  columnNames[which(is.na(columnNames))] = c("Conference", "ConfRecord", 
+                                             "ConfRating", "ConfPower", 
+                                             "ConfOff", "ConfDef", 
+                                             "ConfSoS", "ConfSSF", 
+                                             "ConfEW", "ConfEL")
+  
+  columnNames = paste("massey", columnNames, sep = "")
+  
+  columnNames[columnNames == "masseyTeam"] = "Team"
+  
+  names(ratingsTable) = columnNames
+  
+  return(ratingsTable)
+  
+}

### 2018-09-14 code/functions/predictedWinners.R

@@ -1,0 +1,53 @@
+## Predicted Game Winners #
+
+
+
+predictedWinners = function() {
+  
+  library(rvest)
+  
+  initialHTML = read_html("https://www.oddsshark.com/ncaaf/computer-picks")
+  
+  winners = html_table(initialHTML)
+  
+  winnerShape = lapply(winners, function(x) select(x, ATS))
+  
+  winnerShape = lapply(winnerShape, function(x) data.frame(computerPick = x[2,], 
+                                                           publicPick = x[3,]))
+  
+  winnerShape = data.table::rbindlist(winnerShape)
+  
+  captions = html_nodes(initialHTML, "caption") %>% 
+    html_text
+  
+  winnerShape$game = captions
+  
+  winnerShape$computerPick = gsub("\\s.*", "", winnerShape$computerPick) 
+  
+  winnerShape$publicPick = gsub("\\s.*", "", winnerShape$publicPick)
+  
+  winnerShape$team1 = stringr::str_extract(winnerShape$game, "\\D+(?= Matchup)")
+  
+  winnerShape$team1 = sub("\\s", "", winnerShape$team1)
+  
+  winnerShape$team1Whole = stringr::str_extract(winnerShape$team1, "(?<=\\s).*")
+  
+  winnerShape$team1Abbv = stringr::str_extract(winnerShape$team1, "^\\S+(?=\\s[A-Z][a-z]+)")
+  
+  winnerShape$team2 = stringr::str_extract(winnerShape$game, "(?<=Matchup ).*")
+  
+  winnerShape$team2Whole = stringr::str_extract(winnerShape$team2, "(?<=\\s).*")
+  
+  winnerShape$team2Abbv = stringr::str_extract(winnerShape$team2, "^\\S+(?=\\s[A-Z][a-z]+)")
+  
+  computerPicks = c(winnerShape$team1Whole[winnerShape$computerPick == winnerShape$team1Abbv],
+                    winnerShape$team2Whole[winnerShape$computerPick == winnerShape$team2Abbv])
+  
+  publicPicks = c(winnerShape$team1Whole[winnerShape$publicPick == winnerShape$team1Abbv],
+                  winnerShape$team2Whole[winnerShape$publicPick == winnerShape$team2Abbv])
+  
+  picks = list(computerPicks = computerPicks, 
+               publicPicks = publicPicks)
+  
+  return(picks)
+}

### 2018-09-14 code/reconciliation/shuffle.R

@@ -1,0 +1,47 @@
+######################
+### Data Wrangling ###
+######################
+
+# The following operations were conducted on the complete data
+# so that columns were re-aligned. Given the position of 
+# the Massey variables within the original data, new
+# Massey variables caused disruptions in alignment of 
+# existing columns. To that end, each pull needed to be
+# broken apart, renamed where appropriate, and bound
+# back together.
+
+library(dplyr)
+
+refresh = readr::read_csv("data/deprecated/completeData_deprecated.csv")
+
+refresh = refresh %>% 
+  mutate(marker = ifelse(eventId == "eventId", seq(1:nrow(.)), NA)) %>% 
+  tidyr::fill(., marker, .direction = "down")
+
+markerNames = sort(unique(refresh$marker))
+
+refresh = lapply(markerNames, function(x) {
+  
+  # browser()
+  newDat = refresh %>% 
+    dplyr::filter(marker == x) %>% 
+    select(-marker)
+  
+  for(i in 1:length(newDat)) {
+    if(is.na(newDat[1, i])) {
+     newDat[1, i] = names(newDat)[i] 
+    }
+  }
+  
+  
+  names(newDat) = newDat[1, ]
+  
+  newDat = newDat[-1, ]
+  
+  return(newDat)
+  
+})
+
+refresh = bind_rows(refresh)
+
+write.csv(refresh, "data/deprecated/refreshed.csv", row.names = FALSE)

### 2018-09-14 code/taskSchedule.R

@@ -1,0 +1,16 @@
+################################
+### Scheduling Pull and Prep ###
+################################
+
+library(taskscheduleR)
+
+pullPrepScript = "C:/Users/sberry5/Documents/research/ticketOptions/code/teamOptionsPullPrep.R"
+
+taskscheduler_create(taskname = "optionsPullPrep", rscript = pullPrepScript, 
+                     schedule = "DAILY", starttime = "13:00", 
+                     startdate = format(Sys.Date(), "%m/%d/%Y"))
+
+# WARNING!!
+# This is for the future! Do not uncomment and run until we are done!
+
+# taskscheduler_delete(taskname = "optionsPullPrep")

### 2018-09-14 code/teamOptionsPullPrep.R

@@ -1,0 +1,124 @@
+##################################
+###       Team Options         ### 
+###     Pull, Prep, Write      ###
+##################################
+
+# The following packages will be needed for everything within this file
+# and the sourced functions.
+
+if (!require("pacman")) install.packages("pacman")
+
+pacman::p_load(dplyr, jsonlite, tidyr, readr, rvest, taskscheduleR, data.table)
+
+# All of the functions for this file are sourced from the functions folder. 
+# If you provide the sourceFunction with a character vector for the path on your
+# machine (e.g., "C:/Users/yourName/Documents/work/"), you will be able to run
+# sourceFunction and bring the functions in all at once.
+
+sourceFunction = function(path = NULL) {
+  if(is.null(path)) {
+    source("C:/Users/sberry5/Documents/research/ticketOptions/code/functions/injuryScrape.R")
+    
+    source("C:/Users/sberry5/Documents/research/ticketOptions/code/functions/masseyComps.R")
+    
+    source("C:/Users/sberry5/Documents/research/ticketOptions/code/functions/masseyRatings.R")
+    
+    source("C:/Users/sberry5/Documents/research/ticketOptions/code/functions/predictedWinners.R")
+  } else {
+    source(paste(path, "injuryScrape.R", sep = ""))
+    
+    source(paste(path, "masseyComps.R", sep  = ""))
+    
+    source(paste(path, "masseyRatings.R", sep  = ""))
+    
+    source(paste(path, "predictedWinners.R", sep = ""))
+  }
+}
+
+sourceFunction()
+
+# First, we will get the options data from CFP-RSVP. The information that we want
+# is sitting in a json file.
+
+optionsDat = jsonlite::read_json("https://api.dibitnow.com/api/v1/eventPerformers/getEventPerformers?eventId=1", 
+                                 simplifyVector = TRUE)
+
+# To keep track of date and time for each pull, we will add Sys.time. 
+
+optionsDat$dateTime = Sys.time()
+
+# One of the columns, teamTiers, has a nested list structure. We need to spread
+# it out, collapse it down, and then bind the individual data frames.
+
+spreadData = lapply(1:nrow(optionsDat), function(x) {
+  
+  res = optionsDat$eventPerformerExperiences[[x]] %>% 
+    select(eventPerformerId, 
[…]

### 2018-09-14 dataReconcile.R

@@ -1,29 +1,0 @@
-########################################
-### Data Version and Reconcilliation ###
-########################################
-
-library(dplyr)
-
-# Given the various computer and data source changes, we have a few data files
-# that need checked and bound together.
-
-originalData = readr::read_csv("optionsDat.csv")
-
-# Since the api changed, we need to select key variables, 
-# so that we can get them bound to subsequent files.
-
-originalData = originalData %>% 
-  select(name, dateTime, initialPrice, price, LL_maxPurchaseQuantity, LL_price, 
-         UL_maxPurchaseQuantity, UL_price)
-
-
-updatedData = readr::read_csv("optionsDatNew.csv")
-
-updateV2Data = readr::read_csv("optionsDatV2.csv")
-
-finalData = bind_rows(updateV2Data, updatedData, originalData)
-
-finalData = finalData %>% 
-  mutate(name = gsub("\\.|,", "", name))
-
-write.csv(finalData, "completeData.csv", row.names = FALSE)

### 2018-09-14 injuryScrape.R

@@ -1,40 +1,0 @@
-#####################
-### Injury Scrape ###
-#####################
-
-injuryScrape = function() {
-  
-  library(rvest)
-  
-  injuryPage = read_html("http://www.donbest.com/ncaaf/injuries/")
-  
-  teamNames = html_nodes(injuryPage, "td.statistics_table_header") %>% 
-    html_text()
-  
-  columnNames = html_nodes(injuryPage, "th.statistics_cellrightborder") %>% 
-    html_text() %>% 
-    unique()
-  
-  test = html_table(injuryPage, fill = TRUE) %>% 
-    `[[`(2)
-  
-  test = test[-1, ] %>% 
-    select(X1:X5)
-  
-  names(test) = columnNames
-  
-  test$team = test$Status
-  
-  test$team[!(test$Date %in% teamNames)] = ""
-  
-  test$team[which(test$team == "" & test$Date != "")] = NA
-  
-  test = tidyr::fill(test, team, .direction = "down") %>% 
-    dplyr::filter(Date != "Date", Date != "", 
-           grepl("^[A-Z]", .$Date) != TRUE) %>% 
-    group_by(team) %>% 
-    summarize(n = n())
-  
-  return(test)
-  
-}

### 2018-09-14 masseyComps.R

@@ -1,23 +1,0 @@
-##########################
-### Massey Comparisons ###
-##########################
-
-masseyComps = function() {
-  
-  skipLines = readLines("https://www.masseyratings.com/cf/compare.csv")
-  
-  allComps = readr::read_csv("https://www.masseyratings.com/cf/compare.csv", 
-                             skip = (which(grepl("^Team", skipLines)) - 1))
-  
-  return(allComps)
-  
-}
-
-# library(irr)
-# 
-# iccRatings = allComps %>% 
-#   select(ARG:SOR)
-# 
-# icc(iccRatings, model = "oneway")
-# 
-# kappam.fleiss(iccRatings)

### 2018-09-14 masseyPreviousYear.R

@@ -1,22 +1,0 @@
-############################
-### Massey Previous Year ###
-############################
-
-library(jsonlite)
-
-masseyPage = fromJSON("https://www.masseyratings.com/teamjson.php?t=11604&s=295489", 
-                      flatten = FALSE)
-
-columnNames = masseyPage$CI$title
-
-ratingsTable = matrix(unlist(masseyPage$DI), nrow = length(masseyPage$DI), 
-                      ncol = length(unlist(masseyPage$DI[[1]])), 
-                      byrow = TRUE) %>% 
-  data.frame()
-
-ratingsTable = select(ratingsTable, 
-                      X1, X2)
-
-names(ratingsTable) = c("masseyYearEnd", "Team")
-
-write.csv(ratingsTable, "massey2017YearEnd.csv", row.names = FALSE)

### 2018-09-14 masseyRatings.R

@@ -1,37 +1,0 @@
-######################
-### Massey Ratings ###
-######################
-
-masseyRatings = function() {
-  
-  library(jsonlite)
-  
-  masseyPage = fromJSON("https://www.masseyratings.com/ratejson.php?s=300937&sub=11604", 
-                        flatten = FALSE)
-  
-  columnNames = masseyPage$CI$title
-  
-  ratingsTable = matrix(unlist(masseyPage$DI), nrow = length(masseyPage$DI), 
-                        ncol = length(unlist(masseyPage$DI[[1]])), 
-                        byrow = TRUE) %>% 
-    data.frame()
-  
-  ratingsTable = select(ratingsTable, 
-                        -X2, -X3, 
-                        -X5, -X6)
-  
-  columnNames[which(is.na(columnNames))] = c("Conference", "ConfRecord", 
-                                             "ConfRating", "ConfPower", 
-                                             "ConfOff", "ConfDef", 
-                                             "ConfSoS", "ConfSSF", 
-                                             "ConfEW", "ConfEL")
-  
-  columnNames = paste("massey", columnNames, sep = "")
-  
-  columnNames[columnNames == "masseyTeam"] = "Team"
-  
-  names(ratingsTable) = columnNames
-  
-  return(ratingsTable)
-  
-}

### 2018-09-14 predictedWinners.R

@@ -1,53 +1,0 @@
-## Predicted Game Winners #
-
-
-
-predictedWinners = function(x) {
-  
-  library(rvest)
-  
-  initialHTML = read_html("https://www.oddsshark.com/ncaaf/computer-picks")
-  
-  winners = html_table(initialHTML)
-  
-  winnerShape = lapply(winners, function(x) select(x, ATS))
-  
-  winnerShape = lapply(winnerShape, function(x) data.frame(computerPick = x[2,], 
-                                                           publicPick = x[3,]))
-  
-  winnerShape = data.table::rbindlist(winnerShape)
-  
-  captions = html_nodes(initialHTML, "caption") %>% 
-    html_text
-  
-  winnerShape$game = captions
-  
-  winnerShape$computerPick = gsub("\\s.*", "", winnerShape$computerPick) 
-  
-  winnerShape$publicPick = gsub("\\s.*", "", winnerShape$publicPick)
-  
-  winnerShape$team1 = stringr::str_extract(winnerShape$game, "\\D+(?= Matchup)")
-  
-  winnerShape$team1 = sub("\\s", "", winnerShape$team1)
-  
-  winnerShape$team1Whole = stringr::str_extract(winnerShape$team1, "(?<=\\s).*")
-  
-  winnerShape$team1Abbv = stringr::str_extract(winnerShape$team1, "^\\S+(?=\\s[A-Z][a-z]+)")
-  
-  winnerShape$team2 = stringr::str_extract(winnerShape$game, "(?<=Matchup ).*")
-  
-  winnerShape$team2Whole = stringr::str_extract(winnerShape$team2, "(?<=\\s).*")
-  
-  winnerShape$team2Abbv = stringr::str_extract(winnerShape$team2, "^\\S+(?=\\s[A-Z][a-z]+)")
-  
-  computerPicks = c(winnerShape$team1Whole[winnerShape$computerPick == winnerShape$team1Abbv],
-                    winnerShape$team2Whole[winnerShape$computerPick == winnerShape$team2Abbv])
-  
-  publicPicks = c(winnerShape$team1Whole[winnerShape$publicPick == winnerShape$team1Abbv],
-                  winnerShape$team2Whole[winnerShape$publicPick == winnerShape$team2Abbv])
-  
-  picks = list(computerPicks = computerPicks, 
-               publicPicks = publicPicks)
-  
-  return(picks)
-}

### 2018-09-14 sagarinRatings.R

@@ -1,15 +1,0 @@
-#######################
-### Sagarin Ratings ###
-#######################
-
-sagarinRatings = function() {
-  
-  sagarinPage = read_html("http://sagarin.com/sports/cfsend.htm")
-  
-  sagarinRatings = html_nodes(sagarinPage, "pre") %>% 
-    html_text %>% 
-    `[[`(3)
-  
-  sagarinRows = stringr::str_extract_all(sagarinRatings, "\\r\\n.*\\r\\n")
-  
-}

### 2018-09-14 shuffle.R

@@ -1,47 +1,0 @@
-######################
-### Data Wrangling ###
-######################
-
-# The following operations were conducted on the complete data
-# so that columns were re-aligned. Given the position of 
-# the Massey variables within the original data, new
-# Massey variables caused disruptions in alignment of 
-# existing columns. To that end, each pull needed to be
-# broken apart, renamed where appropriate, and bound
-# back together.
-
-library(dplyr)
-
-completeData = readr::read_csv("completeData_Deprecated.csv")
-
-completeData = completeData %>% 
-  mutate(marker = ifelse(eventId == "eventId", seq(1:nrow(.)), NA)) %>% 
-  tidyr::fill(., marker, .direction = "down")
-
-markerNames = sort(unique(completeData$marker))
-
-refresh = lapply(markerNames, function(x) {
-  
-  # browser()
-  newDat = completeData %>% 
-    dplyr::filter(marker == x) %>% 
-    select(-marker)
-  
-  for(i in 1:length(newDat)) {
-    if(is.na(newDat[1, i])) {
-     newDat[1, i] = names(newDat)[i] 
-    }
-  }
-  
-  
-  names(newDat) = newDat[1, ]
-  
-  newDat = newDat[-1, ]
-  
-  return(newDat)
-  
-})
-
-refresh = bind_rows(refresh)
-
-write.csv(refresh, "refreshed.csv", row.names = FALSE)

### 2018-09-14 taskSchedule.R

@@ -1,16 +1,0 @@
-################################
-### Scheduling Pull and Prep ###
-################################
-
-library(taskscheduleR)
-
-pullPrepScript = "C:/Users/sberry5/Documents/research/ticketOptions/teamOptionsPullPrep.R"
-
-taskscheduler_create(taskname = "optionsPullPrep", rscript = pullPrepScript, 
-                     schedule = "DAILY", starttime = "13:00", 
-                     startdate = format(Sys.Date(), "%m/%d/%Y"))
-
-# WARNING!!
-# This is for the future! Do not uncomment and run until we are done!
-
-# taskscheduler_delete(taskname = "optionsPullPrep")

### 2018-09-14 teamOptionsPullPrep.R

@@ -1,117 +1,0 @@
-##################################
-###       Team Options         ### 
-###     Pull, Prep, Write      ###
-##################################
-
-# The following packages will be needed for everything within this file
-# and the sourced functions.
-
-if (!require("pacman")) install.packages("pacman")
-
-pacman::p_load(dplyr, jsonlite, tidyr, readr, rvest, taskscheduleR)
-
-library(dplyr); library(jsonlite); library(tidyr)
-
-# Sourcing the following files will run the appropriate functions.
-# Before sourcing, however, you will need to specify the appropriate locations
-# for your folder structure. Uncomment the code on lines 20 through 24 to run
-# everything on your own machine.
-
-# pathToMyFolder = "path/to/files/"
-# source(paste(pathToMyFolder, "injuryScrape.R", sep = "))
-# source(paste(pathToMyFolder, "masseyComps.R", sep  = "))
-# source(paste(pathToMyFolder, "masseyRatings.R", sep  = "))
-# source(paste(pathToMyFolder, "predictedWinners.R", sep = "))
-
-
-source("C:/Users/sberry5/Documents/research/ticketOptions/injuryScrape.R")
-
-source("C:/Users/sberry5/Documents/research/ticketOptions/masseyComps.R")
-
-source("C:/Users/sberry5/Documents/research/ticketOptions/masseyRatings.R")
-
-source("C:/Users/sberry5/Documents/research/ticketOptions/predictedWinners.R")
-
-
-# First, we will get the options data from CFP-RSVP. The information that we want
-# is sitting a json file.
-
-optionsDat = jsonlite::read_json("https://api.dibitnow.com/api/v1/eventPerformers/getEventPerformers?eventId=1", 
-                                 simplifyVector = TRUE)
-
-# To keep track of date and time for each pull, we will add Sys.time. 
-
-optionsDat$dateTime = Sys.time()
-
-# One of the columns, teamTiers, has a nested list structure. We need to spread
-# it out, collapse it down, and then bind the individual data frames.
-
-spreadData = lapply(1:nrow(optionsDat), function(x) {
-  
-  res = optionsDat$eventPerformerExperiences[[x]] %>% 
-    select(eventPerformerId, 
-           experienceId, experienceShortName, 
-           initialPrice, maxPurchaseQuantity, price, 
[…]

### 2018-09-14 code/teamOptionsPullPrep.R
updates

@@ -110,9 +110,23 @@ optionsDat = optionsDat %>%
 
 # Before doing anything else, we want to put this data into our runningTab data.
 
-write.table(x = optionsDat, 
-            file = "C:/Users/sberry5/Documents/research/ticketOptions/data/runningTab.csv", append = TRUE, 
-            na = "", sep = ",", row.names = FALSE)
+writeTableInput = function(path = NULL) {
+  if(is.null(path)) {
+    write.table(x = optionsDat, 
+                file = "C:/Users/sberry5/Documents/research/ticketOptions/data/runningTab.csv", append = TRUE, 
+                na = "", sep = ",", row.names = FALSE)
+  } else {
+    write.table(x = optionsDat, 
+                file = paste(path, "data/runningTab.csv", sep = ""), append = TRUE, 
+                na = "", sep = ",", row.names = FALSE)
+  }
+}
+
+# The writeTableInput function behaves the same way as sourceFunctions; just
+# supply the appropriate path and it will go.
+
+writeTableInput()
+
   
 # We want to replace the old past table with the new one, so we will bring in all
 # of the past data, bind, and rewrite.
@@ -121,4 +135,12 @@ pastData = readr::read_csv("C:/Users/sberry5/Documents/research/ticketOptions/da
 
 optionsDat = data.table::rbindlist(list(pastData, optionsDat), fill = TRUE)
 
-write.csv(optionsDat, "C:/Users/sberry5/Documents/research/ticketOptions/data/completeData.csv")
+writeCSVInput = function(path = NULL) {
+  if(is.null(path)) {
+    write.csv(optionsDat, "C:/Users/sberry5/Documents/research/ticketOptions/data/completeData.csv")
+  } else {
+    write.csv(optionsDat, file = paste(path, "data/completeData.csv", sep = ""))
+  }
+}
+
+writeCSVInput()

### 2018-09-14 paper.Rmd
a few paper updates

@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+```{r importPrep}
+pacman::p_load(dplyr, ggplot2)
+
+initialData = readr::read_csv("data/optionsDat.csv")
+
+previousYearRank = readr::read_csv("data/massey2017YearEnd.csv")
+
+previousYearRank = previousYearRank %>% 
+  mutate(Team = as.character(Team), 
+         Team = gsub("St$", "State", Team))
+
+initialData = initialData %>% 
+  mutate(name = gsub("\\.|,", "", name)) %>% 
+  left_join(., previousYearRank, by = c("name" = "Team")) %>% 
+  group_by(name)
+
+top25PrevNames = previousYearRank %>% 
+  arrange(as.numeric(masseyYearEnd)) %>% 
+  head(25) %>% 
+  select(Team)
+
+top25InitialPrices = initialData %>% 
+  slice(., 1L) %>% 
+  dplyr::filter(masseyYearEnd < 26) %>% 
+  select(masseyYearEnd, UL_price, LL_price, name) %>% 
+  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
+  tidyr::gather(key, value, -name, -masseyYearEnd) %>% 
+  arrange(name) %>% 
+  ggplot(., aes(name, as.numeric(value), color = key)) +
+  geom_point(size = 3) +
+  scale_x_discrete(limits = top25PrevNames$Team) +
+  theme_minimal()
+
+completeData = readr::read_csv("refreshed.csv") %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime))
+
+top10Ever = unique(completeData$name[as.numeric(completeData$AP) < 11])
+
+top25Ever = unique(completeData$name[as.numeric(completeData$AP) < 26])
+
+endOfDay = completeData %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
+  group_by(name, day) %>% 
+  slice(n())
+
+volatile = completeData %>% 
+  group_by(name, day) %>%
+  slice(n()) %>%
+  group_by(name) %>% 
+  mutate(lagLL = dplyr::lag(LL_price), 
+         lagUL = dplyr::lag(UL_price)) %>% 
+  mutate(volatilityLL = ((LL_price / lagLL) - 1), 
+         volatilityUL = ((UL_price / lagUL) - 1)) %>% 
+  summarize(sdLL = sd(volatilityLL, na.rm = TRUE), 
+            sdUL = sd(volatilityUL, na.rm = TRUE), 
+            volLL = sdLL * 15.937, 
+            volUL = sdUL * 15.937, 
+            meanLL = mean(LL_price, na.rm = TRUE), 
+            meanUL = mean(UL_price, na.rm = TRUE)) %>% 
+  arrange(desc(volLL), desc(volUL))
 ```
@@
 ```{r}
-
-library(dplyr)
-
-library(ggplot2)
-
-initialData = readr::read_csv("optionsDat.csv")
-
-previousYearRank = readr::read_csv("massey2017YearEnd.csv")
-
-previousYearRank = previousYearRank %>% 
-  mutate(Team = as.character(Team), 
-         Team = gsub("St$", "State", Team))
-
-initialData = initialData %>% 
-  mutate(name = gsub("\\.|,", "", name)) %>% 
-  left_join(., previousYearRank, by = c("name" = "Team")) %>% 
-  group_by(name)
-
-top25PrevNames = previousYearRank %>% 
-  arrange(as.numeric(masseyYearEnd)) %>% 
-  head(25) %>% 
-  select(Team)
-
-initialData %>% 
-  slice(., 1L) %>% 
-  dplyr::filter(masseyYearEnd < 26) %>% 
-  select(masseyYearEnd, UL_price, LL_price, name) %>% 
-  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
-  tidyr::gather(key, value, -name, -masseyYearEnd) %>% 
-  arrange(name) %>% 
-  ggplot(., aes(name, as.numeric(value), color = key)) +
-  geom_point(size = 3) +
-  scale_x_discrete(limits = top25PrevNames$Team) +
-  theme_minimal()
+top25InitialPrices
 ```
@@
 
+Generally, strong correlations are present. For upcoming models, we will opt to use AP, MAS, and USA
+
 ## Replicating Rankings Through Option Prices
@@
 ```{r}
-completeData = readr::read_csv("refreshed.csv") %>% 
-  mutate(day = gsub("\\s.*", "", .$dateTime))
-
-top10Ever = unique(completeData$name[as.numeric(completeData$AP) < 11])
-
-endOfDay = completeData %>% 
-  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
-  group_by(name, day) %>% 
-  slice(n())
-
 completeData %>% 
@@
 Not surprisingly, Ohio State and Oklahoma are starting to trend upwards; meanwhile, Notre Dame has consistently high prices (relatively speaking), but on a potential downward trend.
+
+### Volatility
+
+```{r}
+volatile %>% 
+  filter(name %in% top25Ever) %>% 
+  select(volLL, volUL, name) %>% 
+  tidyr::gather(key = key, value = value, -name) %>% 
+  ggplot(., aes(name, value, color = key)) + 
+  geom_point() +
+  theme_minimal()
+```
 
@@
 
-### Volatility
-
-```{r}
-volatile = completeData %>% 
-  group_by(name, day) %>%
-  slice(n()) %>%
-  group_by(name) %>% 
-  mutate(lagLL = dplyr::lag(LL_price), 
-         lagUL = dplyr::lag(UL_price)) %>% 
-  mutate(volatilityLL = ((LL_price / lagLL) - 1), 
-         volatilityUL = ((UL_price / lagUL) - 1)) %>% 
-  summarize(sdLL = sd(volatilityLL, na.rm = TRUE), 
-            sdUL = sd(volatilityUL, na.rm = TRUE), 
-            volLL = sdLL * 15.937, 
-            volUL = sdUL * 15.937, 
-            meanLL = mean(LL_price, na.rm = TRUE), 
-            meanUL = mean(UL_price, na.rm = TRUE)) %>% 
-  arrange(desc(volLL), desc(volUL))
-```
-
 ###

### 2018-09-14 code/historical.R
historical prices

@@ -1,0 +1,42 @@
+#####################################
+### Historical JSON From CFP-RSVP ###
+#####################################
+
+performerID = readr::read_csv("data/completeData.csv")
+
+ids = performerID$id
+
+rm(performerID)
+
+allHistory = lapply(ids, function(x) {
+  
+  Sys.sleep(.25)
+  
+  pageID = paste("https://api.dibitnow.com/api/v1/eventPerformers/getEventPerformer?eventPerformerId=", 
+                 x, sep = "") 
+  
+  out = jsonlite::read_json(pageID, simplifyVector = TRUE)
+  
+  lowerLevelChange = out$priceTrend$lowerLevelPriceChanges
+  
+  names(lowerLevelChange) = c("lowerLevelPrice", "dateTicks")
+  
+  lowerLevelChange$name = out$name
+  
+  lowerLevelChange = reshape2::melt(lowerLevelChange, 
+                                    id.vars = c("name", "dateTicks"))
+  
+  upperLevelChange = out$priceTrend$upperLevelPriceChanges
+  
+  names(upperLevelChange) = c("upperLevelPrice", "dateTicks")
+  
+  upperLevelChange$name = out$name
+  
+  upperLevelChange = reshape2::melt(upperLevelChange, 
+                                    id.vars = c("name", "dateTicks"))
+  
+  out = bind_rows(lowerLevelChange, upperLevelChange)
+  
+  return(out)
+})
+

### 2018-09-15 paper.Rmd
updates

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
   arrange(desc(volLL), desc(volUL))
+
+allHistory = read.csv("data/allHistory.csv")
 ```
@@
 
+While the intervals in our option price series are not always equivalent, calculating a team's option price volatility can instructive.
+
 ```{r}
@@
   geom_point() +
+  theme_minimal()
+```
+
+We can compare these plotted volatility values with the values that CFP-RSVP uses to graphically illustrate a team's price trend.
+
+First, we can see which 25 teams have had the most price changes throughout the CFP-RSVP's existence.
+Note that the following table displays total price changes (includes both upper and lower level changes).
+
+```{r}
+changes = allHistory %>% 
+  select(name, variable, value) %>% 
+  distinct() %>% 
+  group_by(name) %>% 
+  summarize(n = n()) %>% 
+  arrange(desc(n)) %>% 
+  as.data.frame() %>% 
+  head(25) %>% 
+  mutate(name = factor(name, levels = .$name))
+
+ggplot(changes, aes(name, n)) + 
+  geom_point() + 
+  scale_x_discrete(breaks = changes$name) +
+  theme_minimal()
+```
+
+A cursory glance of the listed names should look familiar.
+
+We can also plot the values:
+
+```{r}
+
+allHistory %>% 
+  filter(allHistory$name %in% top25Ever) %>%
+  group_by(name, variable) %>% 
+  mutate(obs = 1:n()) %>% 
+  ggplot(., aes(obs, value, color = variable)) + 
+  geom_path() +
+  facet_wrap(~name) +
   theme_minimal()
@@
 
-```{r}
+```{r, eval = FALSE}
 library(RQuantLib)

### 2018-09-15 code/historical.R

@@ -4,13 +4,13 @@
 
 performerID = readr::read_csv("data/completeData.csv")
 
-ids = performerID$id
+ids = sort(unique(performerID$id))
 
 rm(performerID)
 
 allHistory = lapply(ids, function(x) {
   
-  Sys.sleep(.25)
+  Sys.sleep(.1)
   
   pageID = paste("https://api.dibitnow.com/api/v1/eventPerformers/getEventPerformer?eventPerformerId=", 
                  x, sep = "") 
@@ -40,3 +40,6 @@ allHistory = lapply(ids, function(x) {
   return(out)
 })
 
+allHistory = data.table::rbindlist(allHistory)
+
+write.csv(allHistory, "data/allHistory.csv", row.names = FALSE)

### 2018-09-15 paper.Rmd
paper updates

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
 ```
@@
 ```{r importPrep}
-pacman::p_load(dplyr, ggplot2)
+pacman::p_load(dplyr, ggplot2, stringr, mgcv, broom)
 
@@
 If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value.
-In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of XXXX for the 2018 Championship game.
+In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of \$2,689 for the 2018 Championship game.
 In financial modeling terms, this system could be considered a call option for speculation.
 When buying an option (the RSVP), the buyer has the right to buy a stock (a ticket) in the future at the strike price (the face value of the ticket).
+
+There is certainly a huge amount of speculation built into this -- at present an option for a lower-level seat for Alabama is very nearly \$600.
+This is a considerable financial risk when purchasing such an option.
+Coupled with a possible face value around \$800, a fan buying an option for \$600 could easily have \$1,400 wrapped up into a ticket.
+This is still less money than what would be spent on the secondary market, so the reward is potentially high.
 
@@
 ```{r}
-top25InitialPrices
+top25InitialPrices + 
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_y_continuous(name = "") +
+  scale_x_discrete(name = "")
 ```
@@
   facet_wrap(~ name) +
-  theme_minimal()
+  theme_minimal() +
+  scale_x_datetime(name = "") +
+  scale_y_continuous(name = "")
 ```
@@
   geom_point() +
-  theme_minimal()
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_y_continuous(name = "Volatility")
 ```
@@
   group_by(name) %>% 
-  summarize(n = n()) %>% 
-  arrange(desc(n)) %>% 
+  summarize(Count = n()) %>% 
+  arrange(desc(Count)) %>% 
   as.data.frame() %>% 
@@
 
-ggplot(changes, aes(name, n)) + 
+ggplot(changes, aes(name, Count)) + 
   geom_point() + 
   scale_x_discrete(breaks = changes$name) +
-  theme_minimal()
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_x_discrete(name = "")
 ```
@@
 
-We can also plot the values:
+We can also plot the series for all of those teams which have appeared in the top 25 this season:
 
 ```{r}
-
 allHistory %>% 
@@
   facet_wrap(~name) +
-  theme_minimal()
+  theme_minimal() +
+  scale_x_continuous(name = "") + 
+  scale_y_continuous(name = "")
 ```
@@
 
-###
+#### Explaining Volatility
 
-```{r, eval = FALSE}
-library(RQuantLib)
+```{r}
+predictors = completeData %>% 
+  select(Mean, n, name, dateTime, masseyRecord) %>% 
+  mutate(wins = as.numeric(stringr::str_extract(masseyRecord, "^[0-9]+"))) %>% 
+  group_by(name) %>% 
+  slice(n())
 
-AmericanOption(type = "call", 
-               underlying = , 
-               strike = 475, 
-               )
+modDat = left_join(volatile, predictors, by = "name")
+
+modLL = gam(volLL ~ s(Mean) + wins + s(n),
+            data = modDat)
+
+# summary(modLL)
+
+knitr::kable(tidy(modLL))
+
+knitr::kable(tidy(modLL, parametric = TRUE))
+
+par(mfrow = c(2, 2))
+
+plot.gam(modLL, all.terms = TRUE)
+
+modUL = gam(volUL ~ s(Mean) + wins + s(n),
+            data = modDat)
+
+# summary(modUL)
+
+knitr::kable(tidy(modUL))
+
+knitr::kable(tidy(modUL, parametric = TRUE))
+
+par(mfrow = c(2, 2))
+
+plot.gam(modUL, all.terms = TRUE)
 ```
+
+#### Explaining Current Prices
+
+```{r}
+currentPriceModDat = completeData %>% 
+  select(Mean, n, name, dateTime, masseyRecord, UL_price, LL_price) %>% 
+  mutate(wins = as.numeric(stringr::str_extract(masseyRecord, "^[0-9]+"))) %>% 
+  group_by(name) %>% 
+  slice(n())
+
+modLL = gam(LL_price ~ s(Mean) + wins + s(n),
+            data = currentPriceModDat)
+
+# summary(modLL)
+
+knitr::kable(tidy(modLL))
+
+knitr::kable(tidy(modLL, parametric = TRUE))
+
+par(mfrow = c(2, 2))
+
+plot.gam(modLL, all.terms = TRUE)
+
+modUL = gam(UL_price ~ s(Mean) + wins + s(n),
+            data = currentPriceModDat)
+
+# summary(modUL)
+
+knitr::kable(tidy(modUL))
+
+knitr::kable(tidy(modUL, parametric = TRUE))
+
+par(mfrow = c(2, 2))
+
+plot.gam(modUL, all.terms = TRUE)
+
+```
+
+# Take Away Messages
+
+While we don't have data back to the inception of CFP-RSVP, we do have a pretty broad swath.
+Given our visualizations and our models, the strategy is clear: buy early and buy towards the top of the projected rankings.
+There is no such thing as parity within college football; of the possible 16 slots combined over the last 4 years, there have been 10 unique teams in the playoffs (with Alabama making 4 appearances, Clemson making 3 appearances, and Ohio State and Oklahoma each having 2 appearances).
+We would likely invest early in the teams that are likely to be around when the playoffs start.

### 2018-09-15 paper.Rmd
updates

@@
 ```{r importPrep}
-pacman::p_load(dplyr, ggplot2, stringr, mgcv, broom)
+pacman::p_load(dplyr, ggplot2, stringr, mgcv, broom, caret)
 
@@
   left_join(., previousYearRank, by = c("name" = "Team")) %>% 
-  group_by(name)
+  group_by(name) %>% 
+  arrange(masseyYearEnd)
 
@@
   dplyr::filter(masseyYearEnd < 26) %>% 
+  ungroup() %>% 
+  mutate(name = factor(name, levels = .$name)) %>% 
   select(masseyYearEnd, UL_price, LL_price, name) %>% 
@@
 top25Ever = unique(completeData$name[as.numeric(completeData$AP) < 26])
+
+currentRankOrder = completeData %>% 
+  group_by(name) %>% 
+  arrange(dateTime) %>% 
+  slice(n()) %>% 
+  select(name, AP) %>% 
+  arrange(AP)
 
@@
 
-# Hypotheses
+# General Research Questions
 
-# Methods
+When examining the option prices, it becomes apparent that the best teams have the highest prices; conversely teams with little chance of making the CFP have very low dollar values.
+Among those top teams, though, what might be driving the prices?
+As a preliminary pass, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics.
+Specifically, we would anticipate that volatility has a positive relationship with wins (i.e., the more wins a team has, the more the option changes), but a negative relationship with rankings (the lower the ranking, the more the option changes).
+Number of injuries, while certainly challenging to a given team, might not be critical at this point within the season.
+Nonetheless, we would anticipate that the number of injuries will drive the option value down.
+We can certainly imagine a world in which strength of schedule has an impact on prices and volatility -- a nightmarish schedule might all but destroy your chances of making it to the playoffs.
 
-## Data
+# Data
 
@@
 
-### Sources
+## Sources
 
-Given the research questions, we sought to collect data from several different sources. Given the excellent compilation of team rankings, Massey's website was used to amass rankings.
+Given the research questions, we sought to collect data from several different sources. Given the excellent compilation of team rankings, Ken Massey's website was used to amass rankings.
 
-Injury counts were pulled from Don Best's website and projected winners were pulled from XXX.
+Injury counts were pulled from Don Best's website and projected winners were pulled from Odds Shark.
 
@@
 
-### Issues and Limitations
+## Issues and Limitations
 
@@
 
-### Previous Year-end Rank and Early Prices
+## Previous Year-end Rank and Early Prices
 
@@
 top25InitialPrices + 
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  scale_y_continuous(name = "") +
-  scale_x_discrete(name = "")
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
+        axis.title.x = element_blank(), 
+        axis.title.y = element_blank())
 ```
 
-The preceding visualization provides a clear picture
+The preceding visualization provides a clear picture of how early prices reflect the previous year rankings.
+This is not terribly surprising, given that the previous year is all of the information informing pricing.
 
@@
 completeData %>% 
-  select(AP, SAG, MAS, USA, FPI) %>% 
+  select(AP, SAG, MAS, USA, FPI, Mean) %>% 
   na.omit() %>% 
@@
 
-Generally, strong correlations are present. For upcoming models, we will opt to use AP, MAS, and USA
+Generally, strong correlations are present.
+For upcoming models, we will opt to use the mean ranking.
+There are a few reasons for this decision: 1) rankings like AP and USA only go up to 25 and 2) we want something that will capture all of the ratings present.
 
-## Replicating Rankings Through Option Prices
-
-### Top 10 Trends
+## Top 10 Trends
 
@@
   mutate(Rank = as.numeric(Rank), 
-         dateTime = lubridate::ymd_hms(dateTime)) %>% 
+         dateTime = lubridate::ymd_hms(dateTime), 
+         name = factor(name, levels = currentRankOrder$name)) %>% 
   dplyr::filter(name %in% top10Ever) %>% 
@@
   theme_minimal() +
-  scale_x_datetime(name = "") +
-  scale_y_continuous(name = "")
+  theme(axis.title.x = element_blank(), 
+        axis.title.y = element_blank())
 ```
@@
 A quick visual inspection of every team that has been in the top 10 this year might suggest persistent high confidence for Alabama, Clemson, and Georgia (but to a slightly less degree).
-Not surprisingly, Ohio State and Oklahoma are starting to trend upwards; meanwhile, Notre Dame has consistently high prices (relatively speaking), but on a potential downward trend.
+Not surprisingly, Ohio State and Oklahoma are starting to trend upwards; meanwhile, Notre Dame has consistently high prices (relative to their real chances of actually making it to the playoffs), but on a potential downward trend.
 
-### Volatility
+## Volatility
 
-While the intervals in our option price series are not always equivalent, calculating a team's option price volatility can instructive.
+While the intervals in our option price series are not always equivalent, calculating a team's option price volatility can be instructive.
 
@@
   filter(name %in% top25Ever) %>% 
+  mutate(name = factor(name, levels = currentRankOrder$name)) %>% 
   select(volLL, volUL, name) %>% 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
+        axis.title.x = element_blank()) +
   scale_y_continuous(name = "Volatility")
 ```
+
+Mississippi State stands out in terms of volatility.
+Desipte finishing around 14 in the 2017-2018 season, their preseason rankings were a bit below that.
+However, they are now beginning to rise in the rankings.
 
@@
 First, we can see which 25 teams have had the most price changes throughout the CFP-RSVP's existence.
-Note that the following table displays total price changes (includes both upper and lower level changes).
+Note that the following visualizations displays the total number of price changes (includes both upper and lower level changes).
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  scale_x_discrete(name = "")
+  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
+        axis.title.x = element_blank())
 ```
 
-A cursory glance of the listed names should look familiar.
+With some exceptions (Texas, Florida State, Arizona, Memphis, Miami, Northwestern, and N.C.
+State), we see that those teams with the most number of changes are well-aligned with the volatility within the teams that have been in the top 25 throughout the season.
 
-We can also plot the series for all of those teams which have appeared in the top 25 this season:
+We can also plot the reported series for all of those teams which have appeared in the top 25 this season:
 
@@
   filter(allHistory$name %in% top25Ever) %>%
+  mutate(name = factor(name, levels = currentRankOrder$name)) %>% 
   group_by(name, variable) %>% 
@@
   theme_minimal() +
-  scale_x_continuous(name = "") + 
-  scale_y_continuous(name = "")
+    theme(axis.title.y = element_blank(), 
+        axis.title.x = element_blank())
 ```
 
-###
-
-```{r}
-library(lme4)
-
-top10 = completeData %>% 
-  # dplyr::filter(name %in% top10Ever) %>% 
-  mutate(Rank = as.numeric(Rank), 
-         dateTime = lubridate::ymd_hms(dateTime), 
-         averagePrice = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
-  select(name, dateTime, averagePrice, Rank, AP, SAG, MAS, USA, 
-         Mean, n, computerPick, publicPick) %>% 
-  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
-  group_by(name, day) %>% 
-  slice(n())
-
-mixedMod = lmer(averagePrice ~ as.numeric(USA) + as.numeric(MAS) + 
-                  (1|name), data = top10)
-```
+Unfortunately, there is no time component to this reported series -- only a series of numbers. So finding out what caused Alabama's dip in the reported numbers would not be possible.
 
@@
 
+We want to explore if current mean ranking (Mean), wins, injury count (n), and strength of schedule (masseySoS) could be predictive of volatility.
+Since we want to allow for as much flexibility in our model as possible, we will use generalized additive models for our models.
+While we will not smooth wins, all other predictors will be smoothed.
+For the sake of space, we will just look at lower level.
+
 ```{r}
 predictors = completeData %>% 
-  select(Mean, n, name, dateTime, masseyRecord) %>% 
+  select(Mean, n, name, dateTime, masseyRecord, masseySoS) %>% 
   mutate(wins = as.numeric(stringr::str_extract(masseyRecord, "^[0-9]+"))) %>% 
@@
 
-modLL = gam(volLL ~ s(Mean) + wins + s(n),
+modLL = gam(volLL ~ s(Mean) + wins + s(n) + s(masseySoS),
             data = modDat)
-
-# summary(modLL)
 
@@
 knitr::kable(tidy(modLL, parametric = TRUE))
+```
 
+For our smoothed terms, we can see that strength of schedule does not do much, but mean ranking and number of injuries do contribute to volatility.
+Mean ranking behaves as anticipated (a near linear negative relationship), but the number of injuries has an interesting relationship with volatility (see the following plot).
+Wins, our lone parametric term, did not have much to contribute.
+
+```{r}
 par(mfrow = c(2, 2))
 
-plot.gam(modLL, all.terms = TRUE)
-
-modUL = gam(volUL ~ s(Mean) + wins + s(n),
-            data = modDat)
-
-# summary(modUL)
-
-knitr::kable(tidy(modUL))
-
-knitr::kable(tidy(modUL, parametric = TRUE))
-
-par(mfrow = c(2, 2))
-
-plot.gam(modUL, all.terms = TRUE)
+plot.gam(modLL, all.terms = TRUE, se = FALSE)
 ```
@@
 
+We will run a similar model, except using current lower level price as our outcome variable.
+
 ```{r}
 currentPriceModDat = completeData %>% 
-  select(Mean, n, name, dateTime, masseyRecord, UL_price, LL_price) %>% 
+  select(Mean, n, name, dateTime, masseyRecord, masseySoS, UL_price, LL_price) %>% 
   mutate(wins = as.numeric(stringr::str_extract(masseyRecord, "^[0-9]+"))) %>% 
@@
 
-modLL = gam(LL_price ~ s(Mean) + wins + s(n),
+modLL = gam(LL_price ~ s(Mean) + wins + s(n)+ s(masseySoS),
             data = currentPriceModDat)
-
-# summary(modLL)
 
@@
 knitr::kable(tidy(modLL, parametric = TRUE))
+```
 
+While wins and strength of schedule did not offer much (again), mean ranking and number of injuries did offer some predictive power.
+The effect of mean ranking on the current price took on an interesting pattern (as seen below), while the number of injuries had a negative relationship (as would be expected).
+
+```{r}
 par(mfrow = c(2, 2))
 
-plot.gam(modLL, all.terms = TRUE)
+plot.gam(modLL, all.terms = TRUE, se = FALSE)
+```
 
-modUL = gam(UL_price ~ s(Mean) + wins + s(n),
-            data = currentPriceModDat)
+## Model Testing
 
-# summary(modUL)
+```{r}
+currentPriceModDat = na.omit(currentPriceModDat)
 
-knitr::kable(tidy(modUL))
+set.seed(1001)
 
-knitr::kable(tidy(modUL, parametric = TRUE))
+trainIndex <- createDataPartition(currentPriceModDat$LL_price, 
+                                  p = .50, 
+                                  list = FALSE, 
+                                  times = 1)
 
-par(mfrow = c(2, 2))
+trainDat = currentPriceModDat[ trainIndex, ]
 
-plot.gam(modUL, all.terms = TRUE)
+testDat  = currentPriceModDat[-trainIndex, ]
 
+trainMod = gam(UL_price ~ s(Mean) + wins + s(n),
+            data = trainDat)
+
+testMod = update(trainMod, data = testDat)
 ```

### 2018-09-15 paper.Rmd
paper updates

@@
 title: "Football Championship Ticket Options"
-output: pdf_document
+header-includes:
+   - \usepackage{float}
+output:
+    pdf_document
+
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.pos = 'h')
 ```
@@
   mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
-  tidyr::gather(key, value, -name, -masseyYearEnd) %>% 
+  tidyr::gather(type, value, -name, -masseyYearEnd) %>% 
   arrange(name) %>% 
-  ggplot(., aes(name, as.numeric(value), color = key)) +
-  geom_point(size = 3) +
-  scale_x_discrete(limits = top25PrevNames$Team) +
-  theme_minimal()
+  ggplot(., aes(name, as.numeric(value), color = type)) +
+  geom_point(size = 3, alpha = .5) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Dark2")
 
-completeData = readr::read_csv("refreshed.csv") %>% 
+completeData = readr::read_csv("data/completeData.csv") %>% 
   mutate(day = gsub("\\s.*", "", .$dateTime))
@@
 
-```{r}
+```{r, fig.height=4}
 top25InitialPrices + 
@@
         axis.title.x = element_blank(), 
-        axis.title.y = element_blank())
+        axis.title.y = element_blank(), 
+        legend.position = "bottom", 
+        legend.title = element_blank())
+
 ```
+
+\
 
@@
 completeData %>% 
-  select(AP, SAG, MAS, USA, FPI, Mean) %>% 
+  select(AP, SAG, MAS, USA, FPI, Mean, Median) %>% 
   na.omit() %>% 
@@
 
-```{r}
-completeData %>% 
+```{r, fig.height=4}
+top10TrendsData = completeData %>% 
   mutate(Rank = as.numeric(Rank), 
@@
   tidyr::gather(key, value, -name, -dateTime) %>% 
-  arrange(name, dateTime) %>% 
-  ggplot(., aes(dateTime, as.numeric(value), color = key, group = key)) + 
+  arrange(name, dateTime)
+
+levels(top10TrendsData$name) = gsub("\\s", "\n", levels(top10TrendsData$name))
+
+ggplot(top10TrendsData, aes(dateTime, as.numeric(value), color = key, group = key)) + 
   geom_path(na.rm = TRUE, linejoin = "mitre") + 
@@
   theme(axis.title.x = element_blank(), 
-        axis.title.y = element_blank())
+        axis.title.y = element_blank(), 
+        legend.position = "bottom", 
+        legend.title = element_blank()) +
+  scale_color_brewer(palette = "Dark2")
 ```
@@
 
-```{r}
+```{r, fig.height=4}
 volatile %>% 
@@
   ggplot(., aes(name, value, color = key)) + 
-  geom_point() +
+  geom_point(alpha = .5) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1), 
-        axis.title.x = element_blank()) +
-  scale_y_continuous(name = "Volatility")
+        axis.title.x = element_blank(), 
+        legend.position = "bottom", 
+        legend.title = element_blank()) +
+  scale_y_continuous(name = "Volatility") +
+  scale_color_brewer(palette = "Dark2")
 ```
@@
 
-```{r}
+```{r, fig.height=4}
 changes = allHistory %>% 
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 1), 
-        axis.title.x = element_blank())
+        axis.title.x = element_blank(), 
+        legend.position = "bottom", 
+        legend.title = element_blank()) +
+  scale_color_brewer(palette = "Dark2")
 ```
@@
 
-```{r}
-allHistory %>% 
+```{r, fig.height=4}
+historyData = allHistory %>% 
   filter(allHistory$name %in% top25Ever) %>%
@@
   group_by(name, variable) %>% 
-  mutate(obs = 1:n()) %>% 
-  ggplot(., aes(obs, value, color = variable)) + 
+  mutate(obs = 1:n())
+
+levels(historyData$name) = gsub("\\s", "\n", levels(historyData$name))
+
+ggplot(historyData, aes(obs, value, color = variable)) + 
   geom_path() +
@@
     theme(axis.title.y = element_blank(), 
-        axis.title.x = element_blank())
+        axis.title.x = element_blank(), 
+        legend.position = "bottom", 
+        legend.title = element_blank()) +
+  scale_color_brewer(palette = "Dark2")
 ```

### 2018-09-16 paper.Rmd
paper updates and code

@@
 ```{r importPrep}
-pacman::p_load(dplyr, ggplot2, stringr, mgcv, broom, caret)
+pacman::p_load(dplyr, ggplot2, stringr, mgcv, broom, caret, GGally)
 
@@
 Instead of a paywall, data is often locked behind terms of service (ToS).
-While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through Javascrip) or ToS.
-With the appropriate accumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS.
+While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or ToS.
+With the appropriate acumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS.
 In scraping our data, we not only read the ToS, but we also examined each site's robots.txt.
@@
 There are also different hotel packages within the data, but our primary focus will be on the two ticket prices.
+For visualizations, both LL_price and UL_price will be presented; for analyses, we will focus our attention for LL_price.
 
@@
 This data source change resulted in a considerable span of data loss.
-Further, issues arised on the automatic nature of data collection.
+Further, issues arose due to the automatic nature of data collection.
 Windows Task Scheduler was used to automatically run our data collection scripts daily -- unfortunately, this will not run if a computer is shut down.
@@
 
-```{r, fig.height=4}
+```{r, fig.height=3.5}
 top25InitialPrices + 
@@
 
-\
-
 The preceding visualization provides a clear picture of how early prices reflect the previous year rankings.
-This is not terribly surprising, given that the previous year is all of the information informing pricing.
+Some teams, like Penn State, Oklahoma, and Ohio State, seem to have started out a bit higher than what the end-of-season ranking would have suggested.
+All of this is not terribly surprising, given that the previous year is all of the information informing pricing.
 
@@
 
-For the sake of exploration and to limit the chances of future multicollinearity, we wanted to explore how some of the more well-known ranking were correlated:
+For the sake of exploration and to limit the chances of future multicollinearity, we wanted to explore how some of the more well-known ranking were correlated.
+We would anticipate that they are generally well correlated, but :
 
@@
   na.omit() %>% 
-  cor() %>% 
-  corrplot::corrplot.mixed()
+  GGally::ggcorr(., label = TRUE, 
+                 low = "#4575b4", mid = "#ffffbf", 
+                 high = "#d73027")
 ```
@@
 For upcoming models, we will opt to use the mean ranking.
-There are a few reasons for this decision: 1) rankings like AP and USA only go up to 25 and 2) we want something that will capture all of the ratings present.
+There are a few reasons for this decision: 1) rankings like AP and USA only go up to 25, 2) there are sometimes significant differences between the more common ratings (AP & USA Today Coaches) and analytic ratings like Massey and Sagarin, and 3) we want something that will capture all of the ratings present.
 
-## Top 10 Trends
+## Top 25 Trends
 
@@
 
-```{r, fig.height=4}
-top10TrendsData = completeData %>% 
+```{r, fig.height=6}
+top25TrendsData = completeData %>% 
   mutate(Rank = as.numeric(Rank), 
@@
          name = factor(name, levels = currentRankOrder$name)) %>% 
-  dplyr::filter(name %in% top10Ever) %>% 
+  dplyr::filter(name %in% top25Ever) %>% 
   select(dateTime, UL_price, LL_price, name) %>% 
@@
 
-levels(top10TrendsData$name) = gsub("\\s", "\n", levels(top10TrendsData$name))
+levels(top25TrendsData$name) = gsub("\\s", "\n", levels(top25TrendsData$name))
 
-ggplot(top10TrendsData, aes(dateTime, as.numeric(value), color = key, group = key)) + 
+ggplot(top25TrendsData, aes(dateTime, as.numeric(value), color = key, group = key)) + 
   geom_path(na.rm = TRUE, linejoin = "mitre") + 
@@
 
-While the intervals in our option price series are not always equivalent, calculating a team's option price volatility can be instructive.
+While the intervals in our option price series are not always equivalent, we do have many time points for each team.
+When dealing with any investment-related data, volatility is usually of interest.
+We can use standard volatility calculations to produce a volatility measures for each team's upper and lower ticket prices.
 
-```{r, fig.height=4}
+```{r, fig.height=3.5}
 volatile %>% 
@@
 Mississippi State stands out in terms of volatility.
-Desipte finishing around 14 in the 2017-2018 season, their preseason rankings were a bit below that.
+Despite finishing around 14 in the 2017-2018 season, their preseason rankings were a bit below that.
 However, they are now beginning to rise in the rankings.
@@
 
-```{r, fig.height=4}
+With some exceptions (Texas, Florida State, Arizona, Memphis, Miami, Northwestern, and N.C.
+State), we see that those teams with the most number of changes are well-aligned with the volatility within the teams that have been in the top 25 throughout the season.
+
+```{r, fig.height=3.2}
 changes = allHistory %>% 
@@
 
-With some exceptions (Texas, Florida State, Arizona, Memphis, Miami, Northwestern, and N.C.
-State), we see that those teams with the most number of changes are well-aligned with the volatility within the teams that have been in the top 25 throughout the season.
-
 We can also plot the reported series for all of those teams which have appeared in the top 25 this season:
 
-```{r, fig.height=4}
+```{r, fig.height=7}
 historyData = allHistory %>% 
@@
 
-Unfortunately, there is no time component to this reported series -- only a series of numbers. So finding out what caused Alabama's dip in the reported numbers would not be possible.
+Unfortunately, there is no time component to this reported series -- only a series of numbers.
+So finding out what caused Alabama's dip in the reported numbers would not be possible.
+Comparing this abbreviated series (essentially a step-like series) to our series with daily recordings is very interesting.
 
@@
 Since we want to allow for as much flexibility in our model as possible, we will use generalized additive models for our models.
-While we will not smooth wins, all other predictors will be smoothed.
-For the sake of space, we will just look at lower level.
+While we will not smooth wins (there really are not enough values at present to effectively smooth it), all other predictors will be smoothed.
+For the sake of space, we will just look at lower level (we do, however, expect that upper and lower levels will behave similarly).
 
@@
 
-modDat = left_join(volatile, predictors, by = "name")
+modDatVol = left_join(volatile, predictors, by = "name")
 
-modLL = gam(volLL ~ s(Mean) + wins + s(n) + s(masseySoS),
-            data = modDat)
+modLLVol = gam(volLL ~ s(Mean) + wins + s(n) + s(masseySoS),
+            data = modDatVol)
 
-knitr::kable(tidy(modLL))
+knitr::kable(tidy(modLLVol))
 
-knitr::kable(tidy(modLL, parametric = TRUE))
+knitr::kable(tidy(modLLVol, parametric = TRUE))
 ```
@@
 For our smoothed terms, we can see that strength of schedule does not do much, but mean ranking and number of injuries do contribute to volatility.
-Mean ranking behaves as anticipated (a near linear negative relationship), but the number of injuries has an interesting relationship with volatility (see the following plot).
-Wins, our lone parametric term, did not have much to contribute.
+Mean ranking behaves as anticipated (a near linear negative relationship), but the number of injuries has an interesting relationship with volatility (see the following plots).
+Wins, our lone parametric term, did not have much to contribute to our model.
 
@@
 
-plot.gam(modLL, all.terms = TRUE, se = FALSE)
+plot.gam(modLLVol, all.terms = TRUE, se = FALSE)
 ```
@@
 
-modLL = gam(LL_price ~ s(Mean) + wins + s(n)+ s(masseySoS),
+modLLPrice = gam(LL_price ~ s(Mean) + wins + s(n)+ s(masseySoS),
             data = currentPriceModDat)
 
-knitr::kable(tidy(modLL))
+knitr::kable(tidy(modLLPrice))
 
-knitr::kable(tidy(modLL, parametric = TRUE))
+knitr::kable(tidy(modLLPrice, parametric = TRUE))
 ```
@@
 
-plot.gam(modLL, all.terms = TRUE, se = FALSE)
+plot.gam(modLLPrice, all.terms = TRUE, se = FALSE)
 ```
@@
 ## Model Testing
+
+In an effort to see how our model predicting lower level volatility might perform in a training/testing scenario, we split our data 50%/50%.
+
+```{r}
+modDatVol = na.omit(modDatVol)
+
+set.seed(1001)
+
+trainIndex = createDataPartition(modDatVol$volLL, 
+                                  p = .50, list = FALSE, times = 1)
+
+trainDatVol = modDatVol[trainIndex, ]
+
+testDatVol  = modDatVol[-trainIndex, ]
+
+trainModVol = train(volLL ~ Mean + wins + n + masseySoS,
+            data = trainDatVol, method = "gam")
+
+testModVol = predict(trainModVol, data = testDatVol)
+```
+
+The following are error statistics from our trained model predicting new data:
+
+```{r}
+knitr::kable(postResample(pred = testModVol, obs = testDatVol$volLL))
+```
+
+Perhaps not the greatest performance between training and testing, but certainly nothing to completely disregard (reasonable $R^2$ given our context and a mean absolute error that isn't embarrassing).
+
+In doing the same 50%/50% split for current price, we have the following results:
 
@@
 
-trainIndex <- createDataPartition(currentPriceModDat$LL_price, 
-                                  p = .50, 
-                                  list = FALSE, 
-                                  times = 1)
+trainIndex = createDataPartition(currentPriceModDat$LL_price, 
+                                  p = .50, list = FALSE, times = 1)
 
-trainDat = currentPriceModDat[ trainIndex, ]
+trainDatPrice = currentPriceModDat[ trainIndex, ]
 
-testDat  = currentPriceModDat[-trainIndex, ]
+testDatPrice  = currentPriceModDat[-trainIndex, ]
 
-trainMod = gam(UL_price ~ s(Mean) + wins + s(n),
-            data = trainDat)
+trainModPrice = train(LL_price ~ Mean + wins + n + masseySoS,
+            data = trainDatPrice, method = "gam")
 
-testMod = update(trainMod, data = testDat)
+testModPrice = predict(trainModPrice, data = testDatPrice)
 ```
 
+```{r}
+knitr::kable(postResample(pred = testModPrice, obs = testDatPrice$LL_price))
+```
+
+These results are not as promising as those generated with our volatility models.
+
 # Take Away Messages
+
+The models presented here are incomplete, given that we are only 3 weeks into the college football season.
+Despite this, we are encouraged that the ticket options are behaving in a predictable manner.
+As data continues to roll in, our models will grow in complexity and predictive power.
+Ultimately, the goal is buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.
 
@@
 There is no such thing as parity within college football; of the possible 16 slots combined over the last 4 years, there have been 10 unique teams in the playoffs (with Alabama making 4 appearances, Clemson making 3 appearances, and Ohio State and Oklahoma each having 2 appearances).
-We would likely invest early in the teams that are likely to be around when the playoffs start.
+We would likely invest early in the teams that are likely to be around when the playoffs start -- even the casual fan can have a pretty good shot at picking the top.

### 2018-09-16 code/paperCode.R

@@ -1,0 +1,237 @@
+if (!require("pacman")) install.packages("pacman")
+
+pacman::p_load(dplyr, ggplot2, stringr, mgcv,
+               broom, caret, GGally, readr, knitr)
+
+initialData = readr::read_csv("data/frozen/optionsDat.csv")
+
+previousYearRank = readr::read_csv("data/frozen/massey2017YearEnd.csv")
+
+allHistory = read.csv("data/frozen/allHistory.csv")
+
+previousYearRank = previousYearRank %>% 
+  mutate(Team = as.character(Team), 
+         Team = gsub("St$", "State", Team))
+
+initialData = initialData %>% 
+  mutate(name = gsub("\\.|,", "", name)) %>% 
+  left_join(., previousYearRank, by = c("name" = "Team")) %>% 
+  group_by(name) %>% 
+  arrange(masseyYearEnd)
+
+top25PrevNames = previousYearRank %>% 
+  arrange(as.numeric(masseyYearEnd)) %>% 
+  head(25) %>% 
+  select(Team)
+
+top25InitialPrices = initialData %>% 
+  slice(., 1L) %>% 
+  dplyr::filter(masseyYearEnd < 26) %>% 
+  ungroup() %>% 
+  mutate(name = factor(name, levels = .$name)) %>% 
+  select(masseyYearEnd, UL_price, LL_price, name) %>% 
+  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
+  tidyr::gather(type, value, -name, -masseyYearEnd) %>% 
+  arrange(name) %>% 
+  ggplot(., aes(name, as.numeric(value), color = type)) +
+  geom_point(size = 3, alpha = .5) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Dark2")
+
+completeData = readr::read_csv("data/frozen/completeData.csv") %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime))
+
+top10Ever = unique(completeData$name[as.numeric(completeData$AP) < 11])
+
+top25Ever = unique(completeData$name[as.numeric(completeData$AP) < 26])
+
+currentRankOrder = completeData %>% 
+  group_by(name) %>% 
+  arrange(dateTime) %>% 
+  slice(n()) %>% 
+  select(name, AP) %>% 
+  arrange(AP)
+
+endOfDay = completeData %>% 
+  mutate(day = gsub("\\s.*", "", .$dateTime)) %>% 
[…]

### 2018-09-16 code/teamOptionsPullPrep.R

@@ -8,7 +8,8 @@
 
 if (!require("pacman")) install.packages("pacman")
 
-pacman::p_load(dplyr, jsonlite, tidyr, readr, rvest, taskscheduleR, data.table)
+pacman::p_load(dplyr, jsonlite, tidyr, readr, rvest, 
+               taskscheduleR, data.table, reshape)
 
 # All of the functions for this file are sourced from the functions folder. 
 # If you provide the sourceFunction with a character vector for the path on your

### 2018-09-16 paper.Rmd
near final paper updates

@@
 ```{r}
-knitr::kable(postResample(pred = testModVol, obs = testDatVol$volLL))
+knitr::kable(postResample(pred = testModVol, obs = testDatVol$volLL), col.names = "")
 ```
@@
 ```{r}
-knitr::kable(postResample(pred = testModPrice, obs = testDatPrice$LL_price))
+knitr::kable(postResample(pred = testModPrice, obs = testDatPrice$LL_price), col.names = "")
 ```

### 2018-10-10 paper.Rmd
updated code

@@
   ungroup() %>% 
-  mutate(name = factor(name, levels = .$name)) %>% 
+  mutate(name = factor(name, levels = top25PrevNames$Team)) %>% 
   select(masseyYearEnd, UL_price, LL_price, name) %>% 

### 2018-10-10 paperFrozen.Rmd
updated code

--- 2018-10-10 paper.Rmd
@@
 
-initialData = readr::read_csv("data/optionsDat.csv")
+initialData = readr::read_csv("data/frozen/optionsDat.csv")
 
-previousYearRank = readr::read_csv("data/massey2017YearEnd.csv")
+previousYearRank = readr::read_csv("data/frozen/massey2017YearEnd.csv")
 
@@
 
-completeData = readr::read_csv("data/completeData.csv") %>% 
+completeData = readr::read_csv("data/frozen/completeData.csv") %>% 
   mutate(day = gsub("\\s.*", "", .$dateTime))
@@
 
-allHistory = read.csv("data/allHistory.csv")
+allHistory = read.csv("data/frozen/allHistory.csv")
 ```
@@
 In capitalizing on the intersection between excitement and fandom, the CFP teamed with OptionIt to introduce an RSVP system for championship game tickets, starting in 2018.
-A fan may by an RSVP (an option) for a team.
+A fan may buy an RSVP (an option) for a team.
 If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value.
@@
 Among those top teams, though, what might be driving the prices?
-As a preliminary pass, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics.
+As a preliminary investigation, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics.
 Specifically, we would anticipate that volatility has a positive relationship with wins (i.e., the more wins a team has, the more the option changes), but a negative relationship with rankings (the lower the ranking, the more the option changes).
@@
 Instead of a paywall, data is often locked behind terms of service (ToS).
-While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or ToS.
+While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or limiting use through restrictive ToS.
 With the appropriate acumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS.
@@
 Since we wanted to conduct our data collection in a programmatic fashion, we wanted to ensure that we were not going to be running afoul of restrictions placed within the robots.txt file.
-To that end, all the data that we collected is free to use -- in other words, no restrictions apply.
+To that end, all the data that we collected is free to use -- in other words, no restrictions apply as far as we know.
 
@@
 
-In an effort to reduce visual clutter and focus on potentially high-value teams, we can visually inspect the top 25 teams at the end of the 2017-2018 season.
+In an effort to reduce visual clutter and focus on potentially high-value teams, we visually inspect the top 25 teams at the end of the 2017-2018 season.
 
-```{r, fig.height=3.5}
+```{r, fig.height=3.25}
 top25InitialPrices + 
@@
 We can use standard volatility calculations to produce a volatility measures for each team's upper and lower ticket prices.
+
+$$ Daily_\Delta = \frac{price - price_{lag}}{price_{lag}} - 1 $$
+$$ volatility = \sigma_{daily_\Delta} * \sqrt[]{252}$$
 
@@
 As data continues to roll in, our models will grow in complexity and predictive power.
-Ultimately, the goal is buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.
+Ultimately, the goal is buy before a team becomes the hot dark horse and sell before a team's dreams burn in flames.
 

### 2018-10-10 code/historical.R

@@ -35,7 +35,7 @@ allHistory = lapply(ids, function(x) {
   upperLevelChange = reshape2::melt(upperLevelChange, 
                                     id.vars = c("name", "dateTicks"))
   
-  out = bind_rows(lowerLevelChange, upperLevelChange)
+  out = dplyr::bind_rows(lowerLevelChange, upperLevelChange)
   
   return(out)
 })

### 2018-10-19 CMSACposter.Rmd
Added CMSACposter.Rmd file and a MCOB image.

--- 2018-10-10 paper.Rmd
@@
 ---
-title: "Football Championship Ticket Options"
-header-includes:
-   - \usepackage{float}
+title: "2018 College Football Playoff Ticket Option Pricing"
+author: "Seth Berry & Scott Nestler; Dept. of IT, Analytics, & Operations"
+date: "10/19/2018"
 output: 
-    pdf_document
-
+  ioslides_presentation:
+    logo: MCOBsquare.jpeg
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.pos = 'h')
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 
-# Introduction
+## Some Background
 
-Since its adoption in the 2014-2015 season, the College Football Playoff (CFP) has given college football fans even more reason to be excited about rankings.
-No longer did a team need to be in the top 2 of the Bowl Championship Series (BCS) to be a championship contender -- instead, the top 4 teams entered into a playoff.
-In capitalizing on the intersection between excitement and fandom, the CFP teamed with OptionIt to introduce an RSVP system for championship game tickets, starting in 2018.
-A fan may by an RSVP (an option) for a team.
-If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value.
-In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of \$2,689 for the 2018 Championship game.
-In financial modeling terms, this system could be considered a call option for speculation.
-When buying an option (the RSVP), the buyer has the right to buy a stock (a ticket) in the future at the strike price (the face value of the ticket).
+- College Football Playoff (CFP) - a 4 team bracket - started with 2014-15 season.
+- In 2018 CFP partnered with OptionIt to initiate a RSVP system for the National Championship Game (and other major bowl games).
+- An RSVP is an **option** - the right (but not an obligation) to buy a ticket for face value, IF your team makes it.
+- Costs (and savings) are high. E.g. for #1-ranked Alabama
+    + An option for a lower-level seat for is nearly \$600. 
+    + With the added face value of roughly \$800, total cost would be \$1,400.
+    + While not cheap, this is likely less than the ticket would cost in the secondary market the week before the game.
 
-There is certainly a huge amount of speculation built into this -- at present an option for a lower-level seat for Alabama is very nearly \$600.
-This is a considerable financial risk when purchasing such an option.
-Coupled with a possible face value around \$800, a fan buying an option for \$600 could easily have \$1,400 wrapped up into a ticket.
-This is still less money than what would be spent on the secondary market, so the reward is potentially high.
+## Questions of Possible Interest
 
-A source of inquiry is what drives the price of these RSVPs.
-Before the season begins, would knowing a teams performance during the previous year drive prices?
-Would preseason ranking have any influence on prices?
-After the start of the season, would rankings, performance, injuries, or other variables affect prices?
-These questions will be explored throughout our paper.
+1. Before the season begins, would knowing a team's performance during the previous year driveprices? 
+2. Would preseason ranking have any influence on prices? 
+3. After the start of the season, would rankings, performance, strengthy of schedule, injuries, or other variables affect prices (and volatility)?
 
-# General Research Questions
+## Previous Year-end Rank & Early Prices
 
-When examining the option prices, it becomes apparent that the best teams have the highest prices; conversely teams with little chance of making the CFP have very low dollar values.
-Among those top teams, though, what might be driving the prices?
-As a preliminary pass, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics.
-Specifically, we would anticipate that volatility has a positive relationship with wins (i.e., the more wins a team has, the more the option changes), but a negative relationship with rankings (the lower the ranking, the more the option changes).
-Number of injuries, while certainly challenging to a given team, might not be critical at this point within the season.
-Nonetheless, we would anticipate that the number of injuries will drive the option value down.
-We can certainly imagine a world in which strength of schedule has an impact on prices and volatility -- a nightmarish schedule might all but destroy your chances of making it to the playoffs.
+- Using Massey's year-end rankings and early option prices (February 2018) we see:
+    + Ranking at end of previous season does generally appear to inform price. 
+    + Some teams started out higher than what might be expected.
 
-# Data
-
-The internet is full of interesting data and finding it is never too hard.
-While finding interesting data does not present a challenge, finding data that is free to use can be tricky.
-In the context of data, free to use does not necessarily mean free in a monetary sense.
-Instead of a paywall, data is often locked behind terms of service (ToS).
-While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or ToS.
-With the appropriate acumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS.
-In scraping our data, we not only read the ToS, but we also examined each site's robots.txt.
-Since we wanted to conduct our data collection in a programmatic fashion, we wanted to ensure that we were not going to be running afoul of restrictions placed within the robots.txt file.
-To that end, all the data that we collected is free to use -- in other words, no restrictions apply.
-
-## Sources
-
-Given the research questions, we sought to collect data from several different sources. Given the excellent compilation of team rankings, Ken Massey's website was used to amass rankings.
-
-Injury counts were pulled from Don Best's website and projected winners were pulled from Odds Shark.
-
-Finally, the RSVP data was pulled from the json object on the CFP-RSVP site.
-While the main page displays the current option price for an upper-level ticket (UL_price), the json object also contains prices for lower-level tickets (LL_price).
-There are also different hotel packages within the data, but our primary focus will be on the two ticket prices.
-For visualizations, both LL_price and UL_price will be presented; for analyses, we will focus our attention for LL_price.
-
-## Issues and Limitations
-
-The RSVP data presented a few challenges along the way -- mainly that the data source changed at some point during the course of data collection.
-This data source change resulted in a considerable span of data loss.
-Further, issues arose due to the automatic nature of data collection.
-Windows Task Scheduler was used to automatically run our data collection scripts daily -- unfortunately, this will not run if a computer is shut down.
-Finally, new ratings were added to the Massey Ratings page.
-Given the position of the Massey ratings in our data join, many columns became displaced from their original position.
-While this was fixed, some values were not recovered.
-
-# Results
-
-## Previous Year-end Rank and Early Prices
-
-As a starting point, we began by looking at Massey's year-end rankings and early option prices (February 2018).
-
-In an effort to reduce visual clutter and focus on potentially high-value teams, we can visually inspect the top 25 teams at the end of the 2017-2018 season.
-
-```{r, fig.height=3.5}
+```{r, fig.height=3}
 top25InitialPrices +
@@
         axis.title.x = element_blank(), 
-        axis.title.y = element_blank(), 
         legend.position = "bottom", 
-        legend.title = element_blank())
+        legend.title = element_blank()) + labs(y = "Price ($)")
 
@@
 
-The preceding visualization provides a clear picture of how early prices reflect the previous year rankings.
-Some teams, like Penn State, Oklahoma, and Ohio State, seem to have started out a bit higher than what the end-of-season ranking would have suggested.
-All of this is not terribly surprising, given that the previous year is all of the information informing pricing.
-
 ## Ranking Correlations
 
-For the sake of exploration and to limit the chances of future multicollinearity, we wanted to explore how some of the more well-known ranking were correlated.
-We would anticipate that they are generally well correlated, but :
+- An exploration of the correlation of well-known rankings was conducted. As expected, there is strong correlation.
+- Going forward, we use the mean ranking for each team among this set.
 
-```{r}
+```{r, fig.height=3.5}
 completeData %>% 
@@
 
-Generally, strong correlations are present.
-For upcoming models, we will opt to use the mean ranking.
-There are a few reasons for this decision: 1) rankings like AP and USA only go up to 25, 2) there are sometimes significant differences between the more common ratings (AP & USA Today Coaches) and analytic ratings like Massey and Sagarin, and 3) we want something that will capture all of the ratings present.
+## Top 10 Trends
 
-## Top 25 Trends
+- The following figure shows the price trend for every team that has been in the AP top 10 since the beginning of the season. 
+- Quick visual inspection suggests persistent high confidence for Alabama, Clemson, and Georgia.
 
-The following figure shows the price trend for every team that has been in the AP top 10 since the beginning of the season.
-
-```{r, fig.height=6}
-top25TrendsData = completeData %>% 
+```{r, fig.height=4}
+top10TrendsData = completeData %>% 
   mutate(Rank = as.numeric(Rank), 
@@
          name = factor(name, levels = currentRankOrder$name)) %>% 
-  dplyr::filter(name %in% top25Ever) %>% 
+  dplyr::filter(name %in% top10Ever) %>% 
   select(dateTime, UL_price, LL_price, name) %>% 
@@
 
-levels(top25TrendsData$name) = gsub("\\s", "\n", levels(top25TrendsData$name))
+levels(top10TrendsData$name) = gsub("\\s", "\n", levels(top10TrendsData$name))
 
-ggplot(top25TrendsData, aes(dateTime, as.numeric(value), color = key, group = key)) + 
+ggplot(top10TrendsData, aes(dateTime, as.numeric(value), color = key, group = key)) + 
   geom_path(na.rm = TRUE, linejoin = "mitre") + 
@@
 
-A quick visual inspection of every team that has been in the top 10 this year might suggest persistent high confidence for Alabama, Clemson, and Georgia (but to a slightly less degree).
-Not surprisingly, Ohio State and Oklahoma are starting to trend upwards; meanwhile, Notre Dame has consistently high prices (relative to their real chances of actually making it to the playoffs), but on a potential downward trend.
-
 ## Volatility
 
-While the intervals in our option price series are not always equivalent, we do have many time points for each team.
-When dealing with any investment-related data, volatility is usually of interest.
-We can use standard volatility calculations to produce a volatility measures for each team's upper and lower ticket prices.
+- The intervals in our option price series are not always equivalent, but we do have many time points for each team. 
+- In financial option pricing,volatility is usually a factor of interest.
 
@@
 
-Mississippi State stands out in terms of volatility.
-Despite finishing around 14 in the 2017-2018 season, their preseason rankings were a bit below that.
-However, they are now beginning to rise in the rankings.
+## Explaining Volatility
 
-We can compare these plotted volatility values with the values that CFP-RSVP uses to graphically illustrate a team's price trend.
-
-First, we can see which 25 teams have had the most price changes throughout the CFP-RSVP's existence.
-Note that the following visualizations displays the total number of price changes (includes both upper and lower level changes).
-
-With some exceptions (Texas, Florida State, Arizona, Memphis, Miami, Northwestern, and N.C.
-State), we see that those teams with the most number of changes are well-aligned with the volatility within the teams that have been in the top 25 throughout the season.
-
-```{r, fig.height=3.2}
-changes = allHistory %>% 
-  select(name, variable, value) %>% 
-  distinct() %>% 
-  group_by(name) %>% 
-  summarize(Count = n()) %>% 
-  arrange(desc(Count)) %>% 
-  as.data.frame() %>% 
-  head(25) %>% 
-  mutate(name = factor(name, levels = .$name))
-
-ggplot(changes, aes(name, Count)) + 
-  geom_point() + 
-  scale_x_discrete(breaks = changes$name) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
-        axis.title.x = element_blank(), 
-        legend.position = "bottom", 
-        legend.title = element_blank()) +
-  scale_color_brewer(palette = "Dark2")
-```
-
-We can also plot the reported series for all of those teams which have appeared in the top 25 this season:
-
-```{r, fig.height=7}
-historyData = allHistory %>% 
-  filter(allHistory$name %in% top25Ever) %>%
-  mutate(name = factor(name, levels = currentRankOrder$name)) %>% 
-  group_by(name, variable) %>% 
-  mutate(obs = 1:n())
-
-levels(historyData$name) = gsub("\\s", "\n", levels(historyData$name))
-
-ggplot(historyData, aes(obs, value, color = variable)) + 
-  geom_path() +
-  facet_wrap(~name) +
-  theme_minimal() +
-    theme(axis.title.y = element_blank(), 
-        axis.title.x = element_blank(), 
-        legend.position = "bottom", 
-        legend.title = element_blank()) +
-  scale_color_brewer(palette = "Dark2")
-```
-
-Unfortunately, there is no time component to this reported series -- only a series of numbers.
-So finding out what caused Alabama's dip in the reported numbers would not be possible.
-Comparing this abbreviated series (essentially a step-like series) to our series with daily recordings is very interesting.
-
-#### Explaining Volatility
-
-We want to explore if current mean ranking (Mean), wins, injury count (n), and strength of schedule (masseySoS) could be predictive of volatility.
-Since we want to allow for as much flexibility in our model as possible, we will use generalized additive models for our models.
-While we will not smooth wins (there really are not enough values at present to effectively smooth it), all other predictors will be smoothed.
-For the sake of space, we will just look at lower level (we do, however, expect that upper and lower levels will behave similarly).
+- Are current mean ranking (Mean), wins, injury count (n), and strength of schedule (masseySoS) predictive of volatility?
+- Using generalized additive models for our models to be as flexible as possible; all parameters but wins are smoothed.
+- Only showing lower level seats here but have similar analysis for upper level.
 
@@
 
-For our smoothed terms, we can see that strength of schedule does not do much, but mean ranking and number of injuries do contribute to volatility.
-Mean ranking behaves as anticipated (a near linear negative relationship), but the number of injuries has an interesting relationship with volatility (see the following plots).
-Wins, our lone parametric term, did not have much to contribute to our model.
+- Strength of schedule and wins did not do much, but mean ranking and number of injuries do contribute to volatility.
+- Mean ranking behaves as anticipated (a near linear negative relationship).
+- Number of injuries has an interesting relationship with volatility (see the following plots).
 
@@
 
-#### Explaining Current Prices
+## Explaining Current Prices
 
@@
 
-# Take Away Messages
+## Take Away Messages
 
-The models presented here are incomplete, given that we are only 3 weeks into the college football season.
-Despite this, we are encouraged that the ticket options are behaving in a predictable manner.
-As data continues to roll in, our models will grow in complexity and predictive power.
-Ultimately, the goal is buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.
-
-While we don't have data back to the inception of CFP-RSVP, we do have a pretty broad swath.
-Given our visualizations and our models, the strategy is clear: buy early and buy towards the top of the projected rankings.
-There is no such thing as parity within college football; of the possible 16 slots combined over the last 4 years, there have been 10 unique teams in the playoffs (with Alabama making 4 appearances, Clemson making 3 appearances, and Ohio State and Oklahoma each having 2 appearances).
-We would likely invest early in the teams that are likely to be around when the playoffs start -- even the casual fan can have a pretty good shot at picking the top.
+- The models presented here are incomplete, as we are only halfway into the college football season. 
+- We are encouraged that the ticket options are behaving in a predictable manner.
+- Our goal is to be able to use these models next season to buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.
+- The general strategy is clear: buy early and buy towards the top of the projected rankings.

### 2018-10-20 CMSACposter.Rmd
setting chunk options to be quiet
Updated slides.

@@
 
-1. Before the season begins, would knowing a team's performance during the previous year driveprices? 
-2. Would preseason ranking have any influence on prices? 
-3. After the start of the season, would rankings, performance, strengthy of schedule, injuries, or other variables affect prices (and volatility)?
+1. Before the season begins, would knowing a team's performance during the previous year drive prices? 
+2. Would initial preseason ranking have any influence on prices? 
+3. After the start of the season, would the following affect prices (and volatility)?
+    - rankings
+    - performance
+    - strength of schedule
+    - injuries
+    - other variables
 
@@
 - Using Massey's year-end rankings and early option prices (February 2018) we see:
-    + Ranking at end of previous season does generally appear to inform price. 
+    + Ranking at end of previous season does *generally* appear to inform price. 
     + Some teams started out higher than what might be expected.
@@
 - The following figure shows the price trend for every team that has been in the AP top 10 since the beginning of the season. 
-- Quick visual inspection suggests persistent high confidence for Alabama, Clemson, and Georgia.
+- Quick visual inspection suggests persistent high confidence for Alabama, Ohio State, and Clemson. In recent weeks, Notre Dame has noticeably increased while Georgia dropped.
 
@@
 
-- The intervals in our option price series are not always equivalent, but we do have many time points for each team. 
-- In financial option pricing,volatility is usually a factor of interest.
+- In financial option pricing, volatility is usually a key factor affecting price.
+- Some schools (Oregon, Wash St, Oklahoma St, USC) have *much* higher volatility than most.
 
@@
 
-## Explaining Volatility
+## Explaining Volatility (for Lower Level)
 
-- Are current mean ranking (Mean), wins, injury count (n), and strength of schedule (masseySoS) predictive of volatility?
-- Using generalized additive models for our models to be as flexible as possible; all parameters but wins are smoothed.
-- Only showing lower level seats here but have similar analysis for upper level.
+- Using generalized additive models to be as flexible as possible; all parameters but wins are smoothed.
 
@@
 
-- Strength of schedule and wins did not do much, but mean ranking and number of injuries do contribute to volatility.
-- Mean ranking behaves as anticipated (a near linear negative relationship).
-- Number of injuries has an interesting relationship with volatility (see the following plots).
+## Explaining Volatility (cont.)
+- Effects of ranking and injuries on volatility seen earlier in season appear to have gone away.
 
@@
 
-While wins and strength of schedule did not offer much (again), mean ranking and number of injuries did offer some predictive power.
-The effect of mean ranking on the current price took on an interesting pattern (as seen below), while the number of injuries had a negative relationship (as would be expected).
+## Explaining Current Prices (cont.)
+
+- Mean ranking appears to offer only significant explanatory power for price.
 
@@
 
-## Model Testing
-
-In an effort to see how our model predicting lower level volatility might perform in a training/testing scenario, we split our data 50%/50%.
-
-```{r}
-modDatVol = na.omit(modDatVol)
-
-set.seed(1001)
-
-trainIndex = createDataPartition(modDatVol$volLL, 
-                                  p = .50, list = FALSE, times = 1)
-
-trainDatVol = modDatVol[trainIndex, ]
-
-testDatVol  = modDatVol[-trainIndex, ]
-
-trainModVol = train(volLL ~ Mean + wins + n + masseySoS,
-            data = trainDatVol, method = "gam")
-
-testModVol = predict(trainModVol, data = testDatVol)
-```
-
-The following are error statistics from our trained model predicting new data:
-
-```{r}
-knitr::kable(postResample(pred = testModVol, obs = testDatVol$volLL), col.names = "")
-```
-
-Perhaps not the greatest performance between training and testing, but certainly nothing to completely disregard (reasonable $R^2$ given our context and a mean absolute error that isn't embarrassing).
-
-In doing the same 50%/50% split for current price, we have the following results:
-
-```{r}
-currentPriceModDat = na.omit(currentPriceModDat)
-
-set.seed(1001)
-
-trainIndex = createDataPartition(currentPriceModDat$LL_price, 
-                                  p = .50, list = FALSE, times = 1)
-
-trainDatPrice = currentPriceModDat[ trainIndex, ]
-
-testDatPrice  = currentPriceModDat[-trainIndex, ]
-
-trainModPrice = train(LL_price ~ Mean + wins + n + masseySoS,
-            data = trainDatPrice, method = "gam")
-
-testModPrice = predict(trainModPrice, data = testDatPrice)
-```
-
-```{r}
-knitr::kable(postResample(pred = testModPrice, obs = testDatPrice$LL_price), col.names = "")
-```
-
-These results are not as promising as those generated with our volatility models.
-
 ## Take Away Messages
@@
 - We are encouraged that the ticket options are behaving in a predictable manner.
-- Our goal is to be able to use these models next season to buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.
-- The general strategy is clear: buy early and buy towards the top of the projected rankings.
+- Our goal is to be able to use these models next season to buy before a team becomes the hot dark horse and sell before a team's dreams burn in flames.
+- The **general strategy** is clear: buy early and buy towards the top of the projected rankings.
